### 2017-01-18 doc/cogsci2017/paper.tex
committing work on cogsci paper

%
% Annual Cognitive Science Conference
% Sample LaTeX Paper -- Proceedings Format
%

% Original : Ashwin Ram (ashwin@cc.gatech.edu)       04/01/1994
% Modified : Johanna Moore (jmoore@cs.pitt.edu)      03/17/1995
% Modified : David Noelle (noelle@ucsd.edu)          03/15/1996
% Modified : Pat Langley (langley@cs.stanford.edu)   01/26/1997
% Latex2e corrections by Ramin Charles Nakisa        01/28/1997
% Modified : Tina Eliassi-Rad (eliassi@cs.wisc.edu)  01/31/1998
% Modified : Trisha Yannuzzi (trisha@ircs.upenn.edu) 12/28/1999 (in process)
% Modified : Mary Ellen Foster (M.E.Foster@ed.ac.uk) 12/11/2000
% Modified : Ken Forbus                              01/23/2004
% Modified : Eli M. Silk (esilk@pitt.edu)            05/24/2005
% Modified: Niels Taatgen (taatgen@cmu.edu) 10/24/2006

%% Change ``a4paper'' in the following line to ``letterpaper'' if you are
%% producing a letter-format document.

\documentclass[10pt,letterpaper]{article}
\setlength{\belowcaptionskip}{5pt plus 3pt minus 2pt}
\usepackage{cogsci}
\usepackage{graphicx}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{pslatex}
\usepackage{apacite}
\usepackage{multirow}
\usepackage{arydshln}
\usepackage{soul,color}
\usepackage{xspace}
%\usepackage[natbib=true]{biblatex}
\usepackage{natbib}
\usepackage{textcomp}
\usepackage[font = small]{caption}
\setlength{\belowcaptionskip}{-7pt}
\def\bibfont{\small}

\title{Present bias in exploratory choice}

\author{  {\large \bf Alexander S. Rich (asr443@nyu.edu)} \\ {\large\bf Todd M. Gureckis (todd.gureckis@nyu.edu)}\\
        New York University, Department of Psychology, 6 Washington Place, New York, NY 10003 USA}
\begin{document}
\maketitle

\begin{abstract}

  \textbf{Keywords} time preference, present bias, exploratory choice
\end{abstract}

The problem of under-exploration seems remarkably common. One of the most severe
expressions of under-exploration is learned helplessness, in which an organism
experiences the absence of contol over the environment, learns that the
environment is uncontrollable, and thus ceases to take actions that might allow
it to regain its sense of control. Learned helplessness has been proposed to
underly most notably depression\cite{Abramson1978, Abramson1989} but also
problems ranging from difficulties in school CITE to poverty CITE.

While the cognitive appraisal of experienced events affects the development of
learned helplessness \citep{Abramson1978}, patterns of exploration clearly play
a role as well \citep{Huys2009, Teodorescu2014a}. In the case of depression,
interventions aimed at increasing the exploration of activities this might
produce a sense of pleasure and mastery have been found to be as effective as
those with a more cognitive orientation \citep{Jacobson1996}.

Under-exploration also seems to occur in the development of complex skills, such
as flying a plane or playing a sport CITE. In these settings, an ``emphasis
change'' training method in which the focus of practice is repeatedly changed
can lead to greater performance gains than unguided practice or more complex
training methods. Emphasis change works because it encourages people to
continually explore the performance space to find more effective cognitive and
motor routines CITE. Without these interventions, people often enter a ``local
maximum'' in which exploration decreases and performance plateaus CITE.

There are many other areas in which under-exploration is less clearly
established, but is suspected to play a role in maladaptive behavior.
Insufficient exploratory interaction with outgroups may be one cause of
stereotypes and prejudice \citep{Denrell2005most}, and interventions that
increase inter-group contact reduce stereotypes \citep{Shook2008}. The crowding
out of exploration by exploitation is a concern in organizational behavior as
well\cite{March1991, Levinthal1993}, prompting research into organizational
structures that may preserve exploration \citep{Fang2010}.

\subsection{Under- and over-exploration in the lab}

Given the seeming pervasiveness of under-exploration in real-world decision
making, one might expect a clear pattern of under-exploration in lab
experiments. Instead, however, lab results are mixed, with people sometimes
under-exploring, sometimes over-exploring, and sometimes exploring close to an
optimal amount. To take two illustrative examples, \citet{Zwick2003} found that
in a sequential search task people under-searched when there were no information
costs but over-searched when there were information costs, and
\citet{Teodorescu2014} found that people explored unknown alternatives too often
or not often enough depending on whether rare outcomes were positive or
negative. Similar results have been obtained within and across a variety of
other studies and paradigms
\citep{tversky1966information, Busemeyer1988, hertwig2004decisions, Navarro2016, Juni2016, Sang2011}.

These experimental studies raise the question of why under-exploration appears
more widespread than over-exploration in the field, but not in the lab. One
possibility is that both forms of deviation from optimality are in fact
prevalent, though perhaps in different settings, and that the seemingly general bias
toward under-exploration is illusory. However, in the current paper we propose
instead that lab tests of exploratory choice have been missing a key component
of real-world exploratory choice, namely, the distribution of choices and
outcomes over time, and that this spreading of exploratory choices over time may
account for the tendency to under-explore.

% In ``observe or bet'' tasks in which people choose to either
% obtain information or reward on each trial, people appear to collect much more
% information than is needed \citep{tversky1966information}. But, in settings
% where the underlying reward probabilities might change, information-seeking
% appears close to optimal \citep{Navarro2016}.

\subsection{Exploration and myopia}

Under-exploration is often proposed to be the result of myopia: the
underweighting of temporally distant rewards relative to close ones \cite{March1991, Levinthal1993}. While
choosing actions that are expected to immediately maximize reward is inherently
exploitative, exploration often requires the decision-maker to suffer short-term
costs in search of better future options.

There are two distinct possible causes of myopia \citep{Bartels2015}. The first is that the decision-maker is
unaware of or fails to consider the future. This form of myopia is likely one
cause of under-exploration, particularly in complex environments. However,
researchers in exploratory choice have begun to address this form of myopia, and
have found that people change their
patterns of exploration based on their expectations about future choices,
showing that, at least to some degree they do consider the future \citep{Meyer1995, Wilson2014a, Rich2017}.

The the second form of myopia is that the decision-maker
is aware of the future but simply doesn't care as much about it.
A wealth have studies have shown that this form of myopia is an important factor
in human decision making, and that people are strongly
biased towards present . For example, people will often prefer a larger, later monetary
reward to a smaller, sooner reward when both rewards are in the future. But when the
time until both rewards is reduced so that the sooner reward is immediate, the
sooner reward seems to loom larger and
people may switch their preference \citep{Kirby1995}. With non-monetary rewards,
such as the cessation of an annoying noise \citep{Solnick1980}, watching a video
when bored \citep{Navarick1998}, or drinking soda
when thirsty \citep{Brown2009}, a bias towards immediate rewards has been observed on the scale of
minutes or even seconds.

Present bias has the clear potential to cause people to explore too little, by
increasing the perceived immediate gains of exploiting relative to the delayed
gains of exploring. However, to date there has been no experimental research
examining how people's bias towards the present affects exploratory choice. In
part, this is a product of limits in experiment design. In most exploratory
choice experiments, each choice and outcome takes fewer than 10 seconds to
complete, making it unlikely for a large bias to develop between the current
trial and future ones. More importantly, regardless of the length of each trial,
rewards are generally not consumed on a trial-by-trial basis. For example, a participant
might recieve ``points'' on each trial, which are converted to money when the
experiment ends. In this sort of paradigm, there is no reason to expect a bias
towards the present choice's rewards over future rewards, as all rewards are
infact delivered at the same time.

In the present study, we seek to address the lack of research on present bias
and exploratory choice. We develop a novel exploratory choice task using
real consumption rewards, and compare behavior when the outcome of each
choice is consumed immediately to behavior when outcomes are temporally delayed.

% why? underexploration is often proposed to be the result of myopia;
% overweighting of the present relative to the future.
% - exploration benefits future, exploitation benefits present.
% - two meanings of myopia (Bartels); one is unaware of the future, or fail to
% consider future, second is aware of the future but simply don't care about it as much.
% - simple adaptive models often exhibit first kind of myopia, and consequently
% underexplore. But experimental studies show that, at least in some cases, people
% are cognizant of the future and adjust their exploration based on it. So unclear
% that unawareness of future is major driver of under-exploration
% - But the second type of myopia, not caring about the future, appears to be very
% common in people. People seems to be present biased

% - this present bias hasn't been studied in exploratory decision making. Most
% studies take place over a short amount of time, such that there are only a few
% seconds dividing the present choice from future choices. Also crucially, rewards
% are usually a medium such as ``points'' that are not consumed immediately, but
% are converted to money all at once at the end of the experiment. Thus present
% and future choices all yield reward at the same time, so a bias towards present
% rewards will likely have no effect on behavior.

% - alternative is that even though exploratory choices contain a present-future
% trade-off, people don't treat them that way, and exploration is driven by things
% like ``exploration bonuses'' that aren't sensitive to time.

\section{Experiment}

\begin{figure*}[htp]
\centering
\includegraphics[width=\textwidth]{figures/machinetimeline.png}
\caption{A schematic illustration of the two conditions. In both conditions, six
30-second work periods pass between consecutive choices with a given machine. In
the immediate condition (left), the outcome for a choice is revealed and occurs immediately after the
choice is made. In the delayed condition (right), the outcome of a choice is revealed
and occurs after a delay of four work periods.}
\label{fig:machinetimeline}
\end{figure*}

\subsection{Method}

\subsubsection{Participants.}

\subsubsection{Design and procedure.}

Participants were informed that their job was to perform a monotonous slider task that
would be split into 30-second ``work periods'', but that they would be able to
make choices throughout the experiment that would give them a chance to watch a
Youtube video instead. The number of remaining work periods in the experiment was shown
at the top of the screen, as was the number of seconds left in the current work period.

The slider task was based of a task previously used by \citet{Gill2012}. In each
period of the slider task, five horizontal sliders appeared on the screen. Each
started at a random setting between 0 and 100, with the slider's value
shown to its right, with a random horizontal offset so that the
sliders were not aligned. The participant's task was to use the mouse to move
each slider to ``50'' before the work period ended. When a participant released
the mouse at the correct setting, the slider turned green to show it had been
completed. To ensure that the task took close to the allotted 30 seconds, at the
beginning of the task only the top slider was enabled, and the other four were
grayed out. Additional sliders were enabled at five-second intervals, such that
all five sliders were available after 20 seconds.

Before beginning the experiment, participants chose one of four videos available
on Youtube: an episode of ``Planet Earth'', and episode of ``The Great British
Bakeoff'', and episode of ``Mythbusters'', or an Ellen Degeneres comedy special.
The video was embedded in the experiment with all user controls (such as
skipping ahead) disabled. When given access to the video, participants had to keep the
browser window open and hold down the space bar for the video to play. This allowed us to ensure that
participants maintained engagement with the content.

In the main part of the experiment, participants completed a total of 70 work
periods. For the first 10 (in the immediate condition) or 6 (in the delayed
condition) work periods, participants simply clicked a button to begin the
slider task. After these initial periods, participants gained access to six
``machines'', each of which had a ``current spinner'' and a ``new spinner''.
Before each work period, participants clicked the circled machine,
were shown the machine and made a choice about which spinner to run. Based on their choices, there was a
chance the machine would perform the work task for the participant, allowing the
participant to watch their chosen video instead.

In the immediate condition, the machine ran immediately after the choice was
made, and affected the next work period. It then ``cooled off'' for the
following five periods, as choices were made with the other five machines.

In the delayed condition, the machine had to ``process'' for four work periods
after a choice was made, thus delaying the outcome by least 2 minutes.
The participant then clicked the machine again to
return to it and observe its outcome, and either perform the slider task or
watch the video. The machine then cooled off for a single period before being ready
for another choice. In the first four periods in which machines were available
in the delayed condition, participants made a choice but then simply pressed a
button to begin the slider task (as no machine had yet finished processing). In
the final four periods, participants observed the outcomes of the machines but
made no more choices.

At each choice, participants had to select between two options: run the
machine's ``current spinner'' or run a ``new spinner''. Each machine had two
circular spinners with arrows at the top. The current spinner was split into
five black and five gold wedges. If a participant chose the
current spinner, it spun (immediately or after four periods, depending
on the condition) and, if the arrow landed on the gold wedge, the machine
worked and the participant could watch the video. Initially, the
current spinner's gold wedges were randomly set for each machine to cover between
1/3 and 2/3 of the spinner.

The new spinner initially showed a question mark. If a participant chose the
new spinner, then (immediately or after four periods) a new spinner was
created and appeared on the machine. The new spinner's gold wedges could cover
anywhere from 0\% of the spinner to 100\% of the spinner. The new spinner then
spun and, if the arrow landed on a gold wedge, the machine worked.
Additionally, if the new spinner had a greater gold area than the current
spinner, the new spinner was ``saved'' and the current spinner was
updated to the new spinner.  This meant that
while choosing a new spinner carried some risk, it also carried long-term
benefits as the current spinner could be improved from its initial value.

Finally, in order to induce exploration throughout the entire experiment, the
six machines would occasionally ``reset'' after they ran. When this occured, a
the current spinner would be set to a new random value between 1/3 and 2/3 gold.
Participants were informed that this would occur randomly on
1/6 of trials. In fact, the procedure was designed so exactly one of the six
machines would reset on each cycle through the machines, and no machine would be
reset on two consecutive uses.

During the machine choices, participants had access to an ``info'' button at the
bottom of the screen that provided reminders about the dynamics of the task. In
addition, before beginning the full experiment, participants completed two
practice phase. First, they performed several trials of practice using the
machines, with the actual work periods removed. Then, they performed three work
periods practicing the slider task.

Participants were given a performance-based bonus of \$3.00 for completing the
slider tasks accurately. If they missed fewer than 10\% of sliders throughout
the experiment and left the video paused less than 20\% of the time, they were
not penalized. However, if they missed more sliders or left the video paused for
longer, they lose 10 cents from their bonus for each additional percentage of
sliders missed or time with the video paused. The running percentage of sliders
missed and video pause time was displayed at the top of the screen throughout
the experiment.

% People often choose actions based not on full information
% about the possible outcomes, but rather on their own past experiences
% \citep{hertwig2004decisions, hertwig2009description}. Making decisions
% from experience is necessary in an uncertain and changing environment,
% but it can cause persistent biases because current beliefs
% and choices can prevent the collection of information that would
% improve future choices. One of the most fundamental biases of
% experience-based decision-making is what \citet{denrell2001adaptation}
% called the ``hot stove effect,'' in which a negative experience with a
% prospect causes an agent to avoid that prospect in the future, preventing
% further belief revision~\citep{denrell2001adaptation,
%   denrell2007adaptive}.  For example, suppose you attend a weekly
% lecture series for the first time and, while the series is usually
% good, you happen to attend a boring talk. This negative experience
% might stop you from attending the series in the future, and as a
% result you might persistently believe the lecture series is boring and
% not attend. This type of false belief can't as easily develop in the
% positive domain; if a lecture series is usually boring but you happen
% to attend a stand-out talk, you're likely to keep attending future
% talks and will soon learn the truth.
% This potential to form false but persistent negative beliefs about
% stochastic prospects, and thus avoid them, has been proposed as a
% possible explanation of risk- and novelty-aversion by people, animals,
% and organizations in a wide variety of contexts
% \citep{denrell2007adaptive, denrell2005most,
%   niv2002evolution}.

% %    
% The hot stove effect is a \emph{learning trap}---a robust
% sub-optimality which follows as a consequence of the incremental
% nature of belief revision \citep{Erev2014}.  In the current paper, we
% describe how the process of selective attention during category
% learning can exacerbate this learning trap, present experimental
% evidence of this bias, and finally propose interventions that may help
% decision makers escape it.

% %The hot stove effect is most typically discussed in the context of
% %``approach-avoid'' decisions, where an agent must repeatedly choose
% %whether to approach an uncertain prospect

% \subsection{Attention and the Hot Stove Effect}

% Past work has focused on the hot stove effect as a problem emerging
% from experience-based decisions about a single stochastic
% prospect which sometimes yields negative outcomes. But real-world 
% environments are more richly structured, with a wide variety of prospects 
% related in complex ways. Rather than mitigate choice biases, such complexity
% may make them worse in ways not considered in past work. We theorize
% that in a complex environment, a pronounced ``attentional'' hot stove
% effect can emerge, even if outcomes are completely deterministic,
% due to people's tendency to categorize the environment based on a low
% number of dimensions or features.

% Most major theories of categorization \citep[e.g.,][]{Nosofsky1986a,
%  love2004sustain} posit that people learn to selectively
% allocate their attention to features that best discriminate category
% members. This conjecture is supported by findings that
% category structures with fewer relevant features are easier to learn
% \citep{Shepard61learning, nosofsky1994comparing} and that people tend 
% to make eye movements only to relevant features \citep{Rehder2005}. 
% In most cases, selective
% attention supports optimal performance by magnifying differences
% between categories. But interestingly, it can also slow learning in
% some cases, particularly when a person learns to \emph{not} attend to
% a certain dimension that may be useful later. This has been observed
% in studies of blocking and backwards blocking \citep{Mackintosh1975,
%   Kruschke2000}, as well as experiments in which subjects were first
% trained on one category structure and then tasked with learning a
% second which involved previously-irrelevant dimensions
% \citep{Kruschke1996}.

% To see how selective attention can ``trap" learners into a persistent
% false belief, consider an environment where an agent encounters
% different prospects that possess or lack each of several features.  Prospects
% that are approached yield a deterministic positive or negative reward 
% but no outcome is experienced when a prospect is avoided.
% The agent's problem now is not just of estimating whether
% a single prospect's value is positive but also of categorization---that
% is, of determining which combinations of features signify that a
% prospect should be approached, and which signify it should be avoided.

% \begin{figure}[t]
% \centering
% \includegraphics[width=\columnwidth]{figures/selectiveattention.pdf}
% \caption{\emph{a:} An deterministic environment containing prospects
%   with two binary features, where only prospects possessing both
%   features are negative. By attending to both features, a
%   decision-maker can avoid all negative prospects while exploiting all
%   positive prospects. \emph{b:} Early experience happens to highlight
%   the relevance of Feature 1. \emph{c:} The agent begins to attend
%   only Feature 1, and ignore Feature 2, making items with and without
%   Feature 2 appear equivalent. Items without Feature 1 are positive,
%   while the value of items with Feature 1 appear stochastic. \emph{d:}
%   The agent now avoids items with Feature 1, since some are
%   negative. This prevents the agent from gaining feedback which would
%   cause it to change its behavior.}
% \label{fig:attention}
% \end{figure}

% Suppose that two features, Feature 1 and Feature 2, are relevant to a
% prospect's value, and only prospects with both features are negative,
% as shown schematically in Figure~\ref{fig:attention}a. As the agent
% begins to gain experience approaching negative prospects
% (Figure~\ref{fig:attention}b), it will likely learn that prospects
% with certain exact combinations of features are negative and should be
% avoided.  But interestingly, it might also try to learn \emph{which}
% of the features was relevant to it being negative.  If experience
% suggests Feature 1 is relevant, for example, it may hypothesize that
% Feature 1 is the sole relevant feature and attend more strongly to
% whether a prospect has Feature 1 in the future. If this tendency is
% extreme, the agent may ignore Feature 2 almost completely
% (Figure~\ref{fig:attention}c).

% If only one dimension is attended, a situation quite like the
% traditional hot stove effect develops. All prospects with Feature 1
% are avoided, including those that lack Feature 2, even though the
% agent had no negative experience with such prospects. What's more, the
% bias away from these prospects is persistent, since avoidance of
% prospects with Feature 1 prevents the agent from collecting the information
% which would cause it to modify its current hypothesis, as shown in
% Figure~\ref{fig:attention}d. The agent may indefinitely
% avoid positive regions of the environment, as well as hold false
% beliefs about how the environment may be divided into meaningful
% categories.

% \subsection{Model simulation}

% To quantitatively verify that biases of the kind described above could
% develop, and their connection to selective attention, we conducted
% several simulations using a version of the ALCOVE model of
% categorization \citep{Kruschke1992a}, modified for a reinforcement
% learning setting in which feedback is action-dependent
% \citep[see][]{Jones2010}. We tested the model on a four-feature
% category structure, where approaching prospects with both Features 1
% and 2 yielded a payoff of $-5$ and approaching any other prospect
% yielded a payoff of $1$. This environment matches the structure depicted in
% Figure~\ref{fig:attention}, but with two added irrelevant features. Simulations were run
% with a specificity constant $c = 6$, a temperature parameter
% $\phi = 15$, an output-weight learning rate $\lambda_w = 0.1$, and an
% attention learning rate $\lambda_\alpha = 0.1$.

% We ran the model for 15 blocks of 16 trials each, across five
% simulation conditions. In the \emph{contingent, att} condition, the
% model only received feedback on the value of a prospect when it
% approached.  In the \emph{full-info, att} condition, the model
% received feedback on the value of all prospects irrespective of
% approach decisions.  The \emph{contingent, no-att} and
% \emph{full-info, no-att} conditions mirrored the first two conditions
% but with the attention-learning parameter was set to zero.  Finally,
% in the \emph{random-info, att} condition the model was yoked to
% receive feedback on the same proportion of trials as the contingent
% model, but the feedback trials were randomly selected and independent
% of the model's choices. The results of these simulations are plotted
% in Figure~\ref{fig:model}.

% \begin{figure*}[htp]
% \centering
% \includegraphics[width=\textwidth]{figures/modelplot.pdf}
% \caption{ALCOVE model simulations of approach-avoid decision-making in
%   five attentional/informational conditions. \emph{Left:} average
%   score per block of 16 trials; \emph{Center:} proportion of positive
%   prospects approached; \emph{Right:} proportion of negative prospects
%   approached. A persistent learning bias develops only when the model
%   received action-dependent feedback and is endowed with selective
%   attention. All conditions were simulated for 1000 model runs.}
% \label{fig:model}
% \end{figure*}

% As depicted in the left-most panel, agents in the contingent
% condition  fell into the learning trap and developed a
% persistent bias, failing to reach perfect performance.  In contrast, 
% agents in the full-info condition quickly reached
% peak performance of 12 points per block. The $p(approach|good)$ and
% $p(approach|bad)$ plots (middle and right panel, respectively) show 
% that the lower performance of the contingent
% condition was not due to continued approach of negative, costly
% prospects, but rather due to the persistent avoidance of some positive
% prospects.

% When attention learning is removed, the persistent bias disappears and
% the model reaches near-peak performance by the last block, though it
% is slower to learn to avoid negative prospects. This should not be
% interpreted simply as evidence that selective attention is generally
% harmful in this environment. On the contrary, with full information,
% selective attention leads to better performance; the full-info,
% no-att condition performs poorly, overgeneralizing from the negative
% prospects. It is only when feedback is dependent on current belief and
% action that selective attention becomes a disadvantage.

% It is also not the case that the contingent condition is biased due
% to a general lack of information. In the random-info condition, where
% the model is given the same number of feedback experiences as the
% contingent condition but distributed evenly over the space of prospects, no
% bias develops. Thus, the attentional hot stove effect occurs due not to an overall
% poverty of information but to a specific pattern of behavior
% which prevents information from being gained about prospects which
% could correct the model's misallocated attention.

% Finally, it is worth noting that in our simulations we assume that
% nothing is encoded in the absence of feedback. However, recent
% experiments have suggested that people might employ constructivist
% coding in the absence of feedback, essentially storing the exemplar as
% though the expected outcome had occurred and reinforcing existing
% beliefs~\citep{henriksson2010coded}. With this
% coding scheme, we would expect the attentional hot stove effect
% exhibited by ALCOVE to be even more pronounced.

% \section{Experiment}

% To test the degree to which people are susceptible to the
% attentional hot stove effect, we performed a simple experiment similar
% to the category-learning task described above.

% \subsection{Method}

% \subsubsection{Participants.}
% One hundred one participants were recruited via Amazon Mechanical
% Turk. Participants received \$1.25 for participation and received a
% performance-based bonus that ranged up to \$1.80.

% \subsubsection{Stimuli.}
% Stimuli were computer-generated cartoon bees that varied on four
% binary dimensions; they had two or six legs, a striped or spotted
% body, single or double wings, and antennae or no antennae, for a total
% of 16 unique stimuli (Figure~\ref{fig:stim}). Two of the four
% dimensions were chosen as relevant, counterbalanced across
% participants. Of the four possible combinations of values on these two
% dimensions, one was chosen at random; stimuli with this combination of
% values were  ``dangerous,'' and the remaining stimuli were
% ``friendly.''

% \subsubsection{Procedure.}
% Participants played the role of a beekeeper collecting honey from
% several beehives. They were told that each hive contained a single
% variety of bees, and that while most hives contained friendly bees
% that would give them honey, some
% hives had been invaded by dangerous bees, which would sting them if
% they tried to harvest.

% In the learning phase, participants encountered each of the 16 bee
% varieties 4 times, for a total of 64 trials. They were informed of the
% number of trials, and a the number of remaining trails was displayed
% throughout learning. Stimuli were ordered such that
% every eight stimuli contained two dangerous and six friendly bee
% varieties, and were otherwise random. On each trial, participants
% visited a new beehive, and were shown one of the bees in the
% hive. Based on the bee's appearance, they then had to choose either to
% attempt to harvest honey from the bee variety in that hive or to avoid
% the hive. When participants chose to harvest, they received honey and
% added \$0.02 to their bonus if the bee variety was friendly, but were
% stung and lost \$0.10 from their bonus if it was dangerous. When
% participants chose to avoid a hive, they gained \$0.00.  Participants
% started the game with a bonus of \$0.40.

% Participants were split into two conditions, which differed in 
% the feedback received upon avoiding a beehive. In the contingent
% condition, no feedback was provided when a participant avoided
% a hive. In the full-info condition, participants still gained
% \$0.00 when they avoided a hive, but were informed of whether the bee
% variety was friendly or dangerous and of what their payoff would have
% been had they harvested.

% The learning phase was followed by a surprise test
% phase.  During the test phase, participants encountered each variety
% twice and chose to harvest or avoid hives as before, but received no
% feedback about the outcomes of their actions and were not able to see
% changes to their bonus. This phase provided a comparison of
% learning under equivalent conditions.

% After the test phase, participants were informed of their total
% bonus, and were asked two final questions: ``About what percentage of
% beehives do you think contained dangerous bees?'' and ``Which features
% do you think were useful in deciding whether a bee variety was
% friendly or dangerous?''. For the first question, participants entered
% a percentage between 0 and 100, and for the
% second question participants could choose any combination of the four
% features using check-boxes.

% \begin{figure}[htp]
% \centering
% \includegraphics[width=\columnwidth]{figures/stimuli.pdf}
% \caption{Example stimuli with opposite values on all four binary dimensions.}
% \label{fig:stim}
% \end{figure}

% \section{Results and Discussion}

% \subsection{Learning}
% Learning performance averaged in 16-trial blocks is shown in Figure~\ref{fig:results},
% \emph{left}, separated by positive and negative prospects. 
% In the first block of learning, participants in the contingent condition
% approached all prospects more than those in the full-info condition,
% $p<.001$.\footnote{All p values are calculated via two-sided permutation test
%   unless otherwise specified.} This suggests participants valued the information which was
% gained by approaching in this condition, in line with the results of
% other recent studies which find that people are information-seeking in
% simple decision-making tasks~\citep{speekenbrink2014uncertainty,
%   rich2014value, Wilson2014a}.

% By the last block of learning participants in both conditions rarely approached
% bad prospects, with no difference between
% the conditions, $p>.25$. Participants in the full-info condition learned
% to approach good prospects at a higher rate by the end of
% learning ($p<.001$), while participants in the contingent condition
% approached good prospects less frequently ($p=.011$), such that by
% the final block participants in the full-info condition were
% significantly more likely to approach a positive prospect than in the
% contingent condition, $p=.017$.

% \begin{figure*}[htp]
% \centering
% \includegraphics[width=\textwidth]{figures/experimentplot.pdf}
% \caption{\emph{Left:} Proportion of negative and positive prospects
%   approached by learning block for the full-info and contingent
%   conditions. Error bars are standard error of the mean. \emph{right:}
%   Performance in the test phase for participants in each condition,
%   plotted as proportion of positive prospects approached against
%   proportion of negative prospects approached and jittered slightly
%   to increase visibility of overlapping points. The point (1,0) represents
%   perfect performance, while the black line denotes chance performance.}
% \label{fig:results}
% \end{figure*}

% \subsection{Test performance}
% Performance in the test phase is plotted in
% Figure~\ref{fig:results}, \emph{right}, which shows each participant's
% proportion of approaching good and bad prospects.
% Participants in the full-info condition were significantly more
% accurate at the test phase, $p=.017$, choosing the correct action on
% 81.5\% of trials versus 71.5\% for the contingent
% condition. Interestingly, they did not gain significantly more points
% on average ($p>.250$), though the difference in median number of
% points scored approached significance ($p=.106$). This lack of
% difference in score is due in part to a small subset of participants
% in the full-info condition who approached all stimuli at a high rate, thus
% incurring a large cost from the bad prospects.

% The higher accuracy of participants in the full-info condition shows
% that they better followed the true, 2-feature rule. However, it does
% not show whether contingent condition participants were less accurate
% because they followed a uni-dimensional rule, or simply because they
% were more noisy. To determine the extent to which participants in each
% condition followed a one-feature rule, we calculated a ``1-feature rule
% score'' for each participant. This score was determined by calculating
% the proportion of trials on which each participant followed each of
% the two relevant 1-feature rules, and then taking the
% maximum over these two proportions. Participants in the full-info
% condition had an average 1-feature rule score of $0.74$, while those
% in the contingent condition had a significantly higher score of
% $0.83$, $p=.004$. Thus, the
% difference in accuracy between the full-info and contingent condition
% participants does not appear to be simply a product of noise due to the
% latter group's restricted information. Rather, it seems that
% participants in the contingent condition systematically attended to only one
% of the two relevant dimensions, thus avoiding a consistent subset of
% rewarding beehives.

% \subsection{Post-test questions}
% Participants in the contingent condition responded on average that
% 35.8\% of prospects were bad, while participants in the full-info
% condition responded that only 28.2\%, a marginally significant
% difference ($p=.055$). The true proportion was 25\%. This supports the conjecture that
% action-dependent feedback can affect a person's beliefs about the
% environment, and is consistent with the findings of
% \citet{fazio2004attitude} that approach-avoid learning lead to belief
% that the environment was more negative than reality. In addition, while 
% only 22.9\% of participants in
% the contingent condition identified the right combination of relevant
% features, 40.4\% of participants in the full-info condition did, a marginally significant difference by Fisher's exact test ($p=.054$).

% In summary, we have provided empirical and computational
% evidence of an attentional
% learning trap wherein avoidance behavior promotes the 
% persistence of false negative beliefs via attentional learning.
% While the overall effect we report is rather intuitive, it is
% important to point out that the vast majority of categorization
% studies have ignored the impact of choice-contingency on
% learning (essentially focusing exclusive on the ``full info" conditions in our
% experiment).  In addition, even the traditional, single prospect
% version of the hot stove effect  has proven
% surprisingly difficult to produce in laboratory
% settings~\citep[e.g.,][]{biele2009learning, rich2014value}.

% \section{Using cognitive models to meliorate the hot stove effect}

% Thinking about choice-contingent learning traps in the context of
% category learning carries several real-world implications. For instance, 
% screening job applicants usually involves a stream of encounters with many 
% different prospects that possess a variety of attributes.  The 
% challenge for a firm is to determine which combinations of attributes tend 
% to signal a good worker.  As in our experiment, it is clear there is
% a strong potential for biased hiring rules that screen out many good 
% candidates due to the selective utilization of certain application features.

% Given these societal concerns, it is interesting to consider if
% insight from computational models of learning can provide guidance on
% how best to limit this bias. When the hot stove effect is caused by
% true stochasticity, it is difficult to prevent
% \citep{denrell2007adaptive}, though considering the future value of
% information can limit its severity \citep{rich2014value}. But in cases
% where the learning trap is primarily attentional, we hypothesize that
% changes to the environment which disrupt the narrowing of attention
%  may significantly reduce biases.  In
% the following section, we explore this issue computationally by asking
% which manipulations to the learning environment help ALCOVE ``avoid
% the trap."

% %Second, it provides a potential avenue to ameliorate the effect,
% %helping people to create more accurate representations of the
% %enviornment and subsequently make better choices. When the hot stove
% %effect is caused by true stochasticity in the environment, it is
% %difficult to prevent. While a proper evaluation of the future value of
% %gaining information can mitigate the effect in situations where it
% %would be most costly \citep{rich2014value}, even an
% %optimal agent exhibits the effect to some degree
% %\citep{denrell2007adaptive}. But when the hot stove effect is due
% %instead to learned inattention to relevant dimensions, there is an
% %increased potential for interventions that could improve long-run
% %belief and choice. Specifically, changes to the environment that
% %disrupt the narrowing of attention onto a small number of dimensions
% %and encourage people to consider a wider set of hypotheses about the
% %environment could make people less likely to maladaptively avoid positive
% %prospects. Here, we briefly explore two interventions that could
% %widen people's attention.

% \subsection{Debiasing interventions}

% In this section, we present two interventions we 
% hypothesized would mitigate the attentional hot
% stove effect.  We first describe the intuitions behind the interventions
% and then provide a modeling analysis testing the effectiveness
% of each.

% \subsubsection{Individuating prospects.}
% One possible way to limit the attentional hot stove effect is to make
% stimuli increasingly distinct and idiosyncratic. When stimuli are more
% distinctive, people tend to treat them more as individuals and show
% increased ability to learn identification compared to
% categorization. While identification learning is more difficult than
% categorization with generic artificial stimuli
% \citep{Shepard61learning, love2004sustain}, \citet{Medin1983} found
% that people were more easily able to pair unique first names than
% categorical last names with photographs of women's
% faces. \citet{love2004sustain} found that this phenomenon could be
% accounted for with the SUSTAIN model of categorization by assuming
% that the women's faces had many distinctive features beyond those
% manipulated by the experimenters, which decreased the similarity among
% stimuli and thus increasing the odds of representing each stimulus
% individually.

% In an approach-avoid decision-making task, increased individuation of stimuli should
% make a person less likely to generalize information gained from
% experience with one prospect to decisions about another. Attention
% paid to idiosyncratic features will slow the biasing of attention
% towards a single dimension, giving the person more opportunity to
% explore a variety of stimuli and learn the true structure of the
% environment. Essentially, increased individuation of prospects shifts
% the task away from category-learning, and towards learning about
% whether to approach individual prospects.  In addition, it may be easier
% for people to track the history of past rewards with more individually
% memorable stimuli.

% \subsubsection{Occluding feature information.}
% A second approach to decreasing the attentional hot stove effect may
% be, paradoxically, to restrict information by randomly occluding some
% features of a prospect such that the decision-maker can't
% observe their values. While this intervention could of course impair a
% person's decision-making ability, it could actually improve
% performance in the long run by causing a greater spread of
% attention. \citet{Taylor2009} found that participants learned more
% about non-diagnostic features in a category-learning task when
% features were randomly occluded, and hypothesized that
% feature-occlusion discourages narrowly selective attention and
% promotes a broader attentional strategy. In the context of approach
% decisions, if a person is attending strongly to a dimension which is
% obscured, he or she may try using other features, which may lead
% to the discovery that they are relevant. Even when the favored
% features is not occluded, the possibility of their future absence may
% cause people to be less quick to look solely at those features
% \citep{Rehder2005}.

% \begin{figure}[htp]
% \centering
% \includegraphics[width=\columnwidth]{figures/interventionplot.pdf}
% \caption{Average performance of ALCOVE model in a test block after four or
%   twenty learning blocks, in the contingent and full-info condition and
%   after two attentional interventions. All conditions simulated for
%   1000 runs of the model.}
% \label{fig:interventions}
% \end{figure}

% \subsection{Modeling debiasing interventions}
% To test the possible efficacy of these interventions, we performed
% model simulations comparing them to the contingent and full-info
% conditions. To modify the model for the individuation condition, we
% added an extra dimension with 16 nominal values representing
% idiosyncratic features of each stimulus
% \citep[following][]{love2004sustain}. For the occluded-dimension condition, on
% $\frac{1}{4}$ of trials a randomly chosen dimension was masked such
% that it did not contribute to the model's network activation and its
% attention weight was not updated.

% The models were simulated for learning phases of two different
% lengths, 4 blocks and 20 blocks, followed by a one block test
% phase with no feedback, where no dimensions were masked in the
% occluded-dimension condition and the individuating dimension was
% masked in the individuation condition. Performance of the models at
% test is reported in Figure~\ref{fig:interventions}.

% With only 4 blocks of learning, as in our
% experiment, the efficacy of the interventions is low. The addition of
% idiosyncratic features aids learning only slightly, and occluding
% dimensions actually hurts performance, as the decrease in information
% from missing dimensions hurts learning more than increasing the spread
% of attention improves it.

% After a more substantial 20 blocks of learning, the interventions
% are more effective. Performance in the individuation condition
% approached that of the full information condition, and performance in
% the occluded-dimension condition surpasses that of the contingent
% condition. Future work will aim to evaluate these interventions with
% human participants, and to develop methods that might speed up their 
% effectiveness such as maximizing the salience of individuating features.

% \section{Conclusion}
% Making decisions from experience is an essential part of
% adaptive cognition, yet such decisions can produce
% biases in action and belief. Here, we considered one
% mechanism through which such biases might develop, which we term the
% attentional hot stove effect.   This effect is a natural consequence of
% popular category learning models but has so far been largely 
% ignored.   To the extent that such biases are caused by misallocation 
% of attention, rather than irreducible stochasticity in the environment, it 
% may be possible to alter decision-making patterns or the environment itself to
% facilitate better learning and choice.

\begin{small}
  \noindent
\textbf{Acknowledgments} 
\end{small}

\bibliographystyle{apacite}
\bibliography{papers}

\end{document}

### 2017-01-26 analysis/get_data.R
adding experiment 1 analysis to paper, committing analysis files

@@ -1,0 +1,82 @@
+library("dplyr")
+## library("RSQLite")
+library("RMySQL")
+library("jsonlite")
+
+## for mysql
+con <- dbConnect(MySQL(),
+         user="lab", password="fishneversink",
+         dbname="mt_experiments", host="gureckislab.org")
+on.exit(dbDisconnect(con))
+rs = dbSendQuery(con, "select * from exploration_consumption")
+data = fetch(rs, n=-1)
+
+## for sqlite
+## my_db <- src_sqlite("../experiment/participants.db")
+## data <- collect(tbl(my_db, "exploration_consumption"))
+
+## data <- data[data$status %in% c(4, 7),]
+data <- data[data$status %in% c(3, 4, 7),]
+data <- data[data$codeversion %in% c("2.4"),]
+
+datastrings <- data$datastring
+datastrings_json <- sapply(datastrings, fromJSON, simplify=F)
+
+trialdata <- sapply(datastrings_json,
+                    function (x) {
+                      y <- x[["data"]][["trialdata"]]
+                      y$templates <- NULL
+                      y$action <- NULL
+                      y$template <- NULL
+                      y$indexOf <- NULL
+                      y$viewTime <- NULL
+                      y;
+                    },
+                    simplify=F)
+
+## trialdata <- datastrings_json[[1]][["data"]][["trialdata"]]
+
+## trialdata$templates <- NULL
+## trialdata$action <- NULL
+## trialdata$template <- NULL
+## trialdata$indexOf <- NULL
+## trialdata$viewTime <- NULL
+names(trialdata) <- NULL
+
+trialdata <- do.call("rbind", trialdata)
+
+
+
+trialdata <- trialdata %>% filter(phase == "EXPERIMENT")
+trialdata$question <- NULL
+trialdata$answer <- NULL
+trialdata$loop <- NULL
+trialdata$status <- NULL
+## trialdata$phase <- NULL
+
+consumptiondata <- trialdata %>% filter(trialType == "consumption")
+exploredata <- trialdata %>% filter(trialType == "exploreexploit")
+
+consumptiondata <- Filter(function(x)!all(is.na(x)), consumptiondata)
+
+exploredata <- Filter(function(x)!all(is.na(x)), exploredata)
+
+
+write.csv(exploredata, file="../data/exploration_data_v2_4.csv")
+write.csv(consumptiondata, file="../data/consumption_data_v2_4.csv")
+
+
+questiondata <- sapply(datastrings_json,
+                    function (x) {
[…]

### 2017-01-26 doc/cogsci2017/paper.tex
adding experiment 1 analysis to paper, committing analysis files

@@
 it to regain its sense of control. Learned helplessness has been proposed to
-underly most notably depression\cite{Abramson1978, Abramson1989} but also
+underly most notably depression \citep{Abramson1978, Abramson1989} but also
 problems ranging from difficulties in school CITE to poverty CITE.
@@
 out of exploration by exploitation is a concern in organizational behavior as
-well\cite{March1991, Levinthal1993}, prompting research into organizational
+well \citep{March1991, Levinthal1993}, prompting research into organizational
 structures that may preserve exploration \citep{Fang2010}.
@@
 Under-exploration is often proposed to be the result of myopia: the
-underweighting of temporally distant rewards relative to close ones \cite{March1991, Levinthal1993}. While
+underweighting of temporally distant rewards relative to close ones \citep{March1991, Levinthal1993}. While
 choosing actions that are expected to immediately maximize reward is inherently
@@
 towards the present choice's rewards over future rewards, as all rewards are
-infact delivered at the same time.
+in fact delivered at the same time.
 
-In the present study, we seek to address the lack of research on present bias
+In the present study, we describe our ongoing efforts to address the lack of research on present bias
 and exploratory choice. We develop a novel exploratory choice task using
-real consumption rewards, and compare behavior when the outcome of each
-choice is consumed immediately to behavior when outcomes are temporally delayed.
+real consumption rewards and present two experiments. In the first, we vary
+the expected payoff of exploration across conditions and test whether this affects
+participant choice, to ensure that participants understand the task and respond
+as predicted to well-studied variables. In the second, we introduce a
+delayed-outcome condition, an intervention only expected to increase exploration
+if people are present biased. While we find no effect of delay on people's
+choices, we discuss plans for future experiments as well as the difficulties of
+studying choice with consumable rewards.
 
-% why? underexploration is often proposed to be the result of myopia;
-% overweighting of the present relative to the future.
-% - exploration benefits future, exploitation benefits present.
-% - two meanings of myopia (Bartels); one is unaware of the future, or fail to
-% consider future, second is aware of the future but simply don't care about it as much.
-% - simple adaptive models often exhibit first kind of myopia, and consequently
-% underexplore. But experimental studies show that, at least in some cases, people
-% are cognizant of the future and adjust their exploration based on it. So unclear
-% that unawareness of future is major driver of under-exploration
-% - But the second type of myopia, not caring about the future, appears to be very
-% common in people. People seems to be present biased
-
-% - this present bias hasn't been studied in exploratory decision making. Most
-% studies take place over a short amount of time, such that there are only a few
-% seconds dividing the present choice from future choices. Also crucially, rewards
-% are usually a medium such as ``points'' that are not consumed immediately, but
-% are converted to money all at once at the end of the experiment. Thus present
-% and future choices all yield reward at the same time, so a bias towards present
-% rewards will likely have no effect on behavior.
-
-% - alternative is that even though exploratory choices contain a present-future
-% trade-off, people don't treat them that way, and exploration is driven by things
-% like ``exploration bonuses'' that aren't sensitive to time.
-
-\section{Experiment}
-
-\begin{figure*}[htp]
-\centering
-\includegraphics[width=\textwidth]{figures/machinetimeline.png}
-\caption{A schematic illustration of the two conditions. In both conditions, six
-30-second work periods pass between consecutive choices with a given machine. In
-the immediate condition (left), the outcome for a choice is revealed and occurs immediately after the
-choice is made. In the delayed condition (right), the outcome of a choice is revealed
-and occurs after a delay of four work periods.}
-\label{fig:machinetimeline}
-\end{figure*}
+\section{Experiment 1}
 
@@
 
+Thirty-one participants completed the experiment, which was conducted over
+Amazon Mechanical Turk using the psiTurk framework \citep{Gureckis2015a}.
+Participants were paid \$5.00 for their participation, with a performance-based
+bonus of up to \$3.00. All participants received the full \$3.00 bonus.
+
 \subsubsection{Design and procedure.}
+
+\begin{figure*}[htp]
+\centering
+\includegraphics[width=\textwidth]{figures/machinetimeline.png}
+\caption{A schematic illustration of the experiment timeline. In all experiments
+  and conditions, six 30-second work periods pass between consecutive choices
+  with a given machine. In Experiment 1 and the immediate condition (left) of
+  Experiment 2, the outcome for a choice is revealed and occurs immediately
+  after the choice is made. In the delayed condition of Experiment 2 (right),
+  the outcome of a choice is revealed and occurs after a delay of four work
+  periods.}
+\label{fig:machinetimeline}
+\end{figure*}
 
@@
 In the main part of the experiment, participants completed a total of 70 work
-periods. For the first 10 (in the immediate condition) or 6 (in the delayed
-condition) work periods, participants simply clicked a button to begin the
+periods. For the first 10 work periods, participants simply clicked a button to begin the
 slider task. After these initial periods, participants gained access to six
@@
 
-In the immediate condition, the machine ran immediately after the choice was
-made, and affected the next work period. It then ``cooled off'' for the
+The machine ran immediately after the choice was
+made, and affected the next work period, as shown in
+Figure~\ref{fig:machinetimeline} (left). It then ``cooled off'' for the
 following five periods, as choices were made with the other five machines.
-
-In the delayed condition, the machine had to ``process'' for four work periods
-after a choice was made, thus delaying the outcome by least 2 minutes.
-The participant then clicked the machine again to
-return to it and observe its outcome, and either perform the slider task or
-watch the video. The machine then cooled off for a single period before being ready
-for another choice. In the first four periods in which machines were available
-in the delayed condition, participants made a choice but then simply pressed a
-button to begin the slider task (as no machine had yet finished processing). In
-the final four periods, participants observed the outcomes of the machines but
-made no more choices.
 
@@
 new spinner, then (immediately or after four periods) a new spinner was
-created and appeared on the machine. The new spinner's gold wedges could cover
-anywhere from 0\% of the spinner to 100\% of the spinner. The new spinner then
+created and appeared on the machine. In the high-explore condition, the new spinner's gold wedges could cover
+anywhere from 25\% of the spinner to 100\% of the spinner. In the low-explore condition, the new spinner's gold wedges could cover
+anywhere from 0\% of the spinner to 75\% of the spinner. The new spinner then
 spun and, if the arrow landed on a gold wedge, the machine worked.
@@
 updated to the new spinner.  This meant that
-while choosing a new spinner carried some risk, it also carried long-term
+while choosing a new spinner carried risk, it could carry long-term
 benefits as the current spinner could be improved from its initial value.
+Importantly, the value of choosing a new spinner was higher in the high-spinner
+condition, because the range of outcomes was shifted towards higher values.
 
@@
 
-% People often choose actions based not on full information
-% about the possible outcomes, but rather on their own past experiences
-% \citep{hertwig2004decisions, hertwig2009description}. Making decisions
-% from experience is necessary in an uncertain and changing environment,
-% but it can cause persistent biases because current beliefs
-% and choices can prevent the collection of information that would
-% improve future choices. One of the most fundamental biases of
-% experience-based decision-making is what \citet{denrell2001adaptation}
-% called the ``hot stove effect,'' in which a negative experience with a
-% prospect causes an agent to avoid that prospect in the future, preventing
-% further belief revision~\citep{denrell2001adaptation,
-%   denrell2007adaptive}.  For example, suppose you attend a weekly
-% lecture series for the first time and, while the series is usually
-% good, you happen to attend a boring talk. This negative experience
-% might stop you from attending the series in the future, and as a
-% result you might persistently believe the lecture series is boring and
-% not attend. This type of false belief can't as easily develop in the
-% positive domain; if a lecture series is usually boring but you happen
-% to attend a stand-out talk, you're likely to keep attending future
-% talks and will soon learn the truth.
-% This potential to form false but persistent negative beliefs about
-% stochastic prospects, and thus avoid them, has been proposed as a
-% possible explanation of risk- and novelty-aversion by people, animals,
-% and organizations in a wide variety of contexts
-% \citep{denrell2007adaptive, denrell2005most,
-%   niv2002evolution}.
+Following the experiment, participants were asked to rate their enjoyment of the
+slider task and of the video-watching task on a 1 to 7 scale.
 
-% %    
-% The hot stove effect is a \emph{learning trap}---a robust
-% sub-optimality which follows as a consequence of the incremental
-% nature of belief revision \citep{Erev2014}.  In the current paper, we
-% describe how the process of selective attention during category
-% learning can exacerbate this learning trap, present experimental
-% evidence of this bias, and finally propose interventions that may help
-% decision makers escape it.
+\section{Results}
 
-% %The hot stove effect is most typically discussed in the context of
-% %``approach-avoid'' decisions, where an agent must repeatedly choose
-% %whether to approach an uncertain prospect
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{figures/exp1results.pdf}
+\caption{Model-based estimates of participants' probability of choosing a new
+  spinner for different values of the current spinner in Experiment 1. Thick
+  lines and shaded regions indicate the mean and 95\% posterior interval for the
+  population-level parameters, while the thin lines indicate the mean posterior
+  parameters for each of the 31 individual participants. Participants in the
+  high-explore condition were more
+  likely to explore at a given current-spinner value than those in the
+  low-explore condition.}
+\label{fig:exp1results}
+\end{figure}
 
-% \subsection{Attention and the Hot Stove Effect}
+Participants rated their enjoyment of the video at 5.77 out of 7, on average,
+and their enjoyment of the slider task at 2.87 out of 7, on average. This
+difference in enjoyment between the tasks was significant, $t(30)=10.53$, $p<.001$.
 
-% Past work has focused on the hot stove effect as a problem emerging
-% from experience-based decisions about a single stochastic
-% prospect which sometimes yields negative outcomes. But real-world 
-% environments are more richly structured, with a wide variety of prospects 
-% related in complex ways. Rather than mitigate choice biases, such complexity
-% may make them worse in ways not considered in past work. We theorize
-% that in a complex environment, a pronounced ``attentional'' hot stove
-% effect can emerge, even if outcomes are completely deterministic,
-% due to people's tendency to categorize the environment based on a low
-% number of dimensions or features.
+To analyze participants trial-by-trial decision-making behavior, we conducted a hierarchical
+Bayesian logistic regression using the Stan modeling language \citep{stan-software:2015}.
+This approach allowed us to estimate population-level effects of the
+current-spinner value and of condition, while also allowing for individual
+differences. The regression model included an intercept term as well as terms
+for the value of the current spinner, the participant's condition, and a
+condition by value of current spinner interaction.
 
-% Most major theories of categorization \citep[e.g.,][]{Nosofsky1986a,
-%  love2004sustain} posit that people learn to selectively
-% allocate their attention to features that best discriminate category
-% members. This conjecture is supported by findings that
-% category structures with fewer relevant features are easier to learn
-% \citep{Shepard61learning, nosofsky1994comparing} and that people tend 
-% to make eye movements only to relevant features \citep{Rehder2005}. 
-% In most cases, selective
-% attention supports optimal performance by magnifying differences
-% between categories. But interestingly, it can also slow learning in
-% some cases, particularly when a person learns to \emph{not} attend to
-% a certain dimension that may be useful later. This has been observed
-% in studies of blocking and backwards blocking \citep{Mackintosh1975,
-%   Kruschke2000}, as well as experiments in which subjects were first
-% trained on one category structure and then tasked with learning a
-% second which involved previously-irrelevant dimensions
-% \citep{Kruschke1996}.
+The results of the regression are shown in Figure~\ref{fig:exp1results}.
+Because we observed few trails for each participant, and the current-spinner
+values observed for each participant depend on chance and the participant's
+choices, we present individual-level model fits rather than the raw choice data.
+We find that at the population level, participants are less likely to choose a
+new spinner when the value of the current spinner is higher, with posterior
+probability $p>.999$. We also find that participants in the high-explore are more likely to explore a new spinner for a
+given current spinner value than those in the low-explore condition, with
+posterior probability $p>.99$. We found no interaction between condition and the
+effect of current-spinner value.
 
-% To see how selective attention can ``trap" learners into a persistent
-% false belief, consider an environment where an agent encounters
-% different prospects that possess or lack each of several features.  Prospects
-% that are approached yield a deterministic positive or negative reward 
-% but no outcome is experienced when a prospect is avoided.
-% The agent's problem now is not just of estimating whether
-% a single prospect's value is positive but also of categorization---that
-% is, of determining which combinations of features signify that a
-% prospect should be approached, and which signify it should be avoided.
+\section{Experiment 2}
 
-% \begin{figure}[t]
-% \centering
-% \includegraphics[width=\columnwidth]{figures/selectiveattention.pdf}
-% \caption{\emph{a:} An deterministic environment containing prospects
-%   with two binary features, where only prospects possessing both
-%   features are negative. By attending to both features, a
-%   decision-maker can avoid all negative prospects while exploiting all
-%   positive prospects. \emph{b:} Early experience happens to highlight
-%   the relevance of Feature 1. \emph{c:} The agent begins to attend
-%   only Feature 1, and ignore Feature 2, making items with and without
-%   Feature 2 appear equivalent. Items without Feature 1 are positive,
-%   while the value of items with Feature 1 appear stochastic. \emph{d:}
-%   The agent now avoids items with Feature 1, since some are
-%   negative. This prevents the agent from gaining feedback which would
-%   cause it to change its behavior.}
-% \label{fig:attention}
-% \end{figure}
+\subsection{Method}
 
-% Suppose that two features, Feature 1 and Feature 2, are relevant to a
-% prospect's value, and only prospects with both features are negative,
-% as shown schematically in Figure~\ref{fig:attention}a. As the agent
-% begins to gain experience approaching negative prospects
-% (Figure~\ref{fig:attention}b), it will likely learn that prospects
-% with certain exact combinations of features are negative and should be
-% avoided.  But interestingly, it might also try to learn \emph{which}
-% of the features was relevant to it being negative.  If experience
-% suggests Feature 1 is relevant, for example, it may hypothesize that
-% Feature 1 is the sole relevant feature and attend more strongly to
-% whether a prospect has Feature 1 in the future. If this tendency is
-% extreme, the agent may ignore Feature 2 almost completely
-% (Figure~\ref{fig:attention}c).
+\subsubsection{Participants.}
 
-% If only one dimension is attended, a situation quite like the
-% traditional hot stove effect develops. All prospects with Feature 1
-% are avoided, including those that lack Feature 2, even though the
-% agent had no negative experience with such prospects. What's more, the
-% bias away from these prospects is persistent, since avoidance of
-% prospects with Feature 1 prevents the agent from collecting the information
-% which would cause it to modify its current hypothesis, as shown in
-% Figure~\ref{fig:attention}d. The agent may indefinitely
-% avoid positive regions of the environment, as well as hold false
-% beliefs about how the environment may be divided into meaningful
-% categories.
+\subsubsection{Design and procedure.}
 
-% \subsection{Model simulation}
+Experiment 2 followed the design of Experiment 1, with the following changes. In
+both conditions, the possible winning proportion of a new spinner ranged fro 0\% to 100\% of
+the spinner. In the immediate condition the machine ran immediately after the
+choice was made, as in Experiment 1. In the delayed condition, each machine was
+presented to the participant four work periods before it was schedules to run,
+and the participant made a choice at that time. The machine then had to
+``process'' for four work periods, thus delaying
+the outcome by over 2 minutes. The participant then clicked the machine again to
+return to it and observe its outcome, and either perform the slider task or
+watch the video. The machine then cooled of for a single period before being
+ready for another choice. The timeline for the delayed condition is shown in
+Figure~\ref{fig:machinetimeline} (right).
 
-% To quantitatively verify that biases of the kind described above could
-% develop, and their connection to selective attention, we conducted
-% several simulations using a version of the ALCOVE model of
-% categorization \citep{Kruschke1992a}, modified for a reinforcement
-% learning setting in which feedback is action-dependent
-% \citep[see][]{Jones2010}. We tested the model on a four-feature
-% category structure, where approaching prospects with both Features 1
-% and 2 yielded a payoff of $-5$ and approaching any other prospect
-% yielded a payoff of $1$. This environment matches the structure depicted in
-% Figure~\ref{fig:attention}, but with two added irrelevant features. Simulations were run
-% with a specificity constant $c = 6$, a temperature parameter
-% $\phi = 15$, an output-weight learning rate $\lambda_w = 0.1$, and an
-% attention learning rate $\lambda_\alpha = 0.1$.
+\section{results}
 
-% We ran the model for 15 blocks of 16 trials each, across five
-% simulation conditions. In the \emph{contingent, att} condition, the
-% model only received feedback on the value of a prospect when it
-% approached.  In the \emph{full-info, att} condition, the model
-% received feedback on the value of all prospects irrespective of
-% approach decisions.  The \emph{contingent, no-att} and
-% \emph{full-info, no-att} conditions mirrored the first two conditions
-% but with the attention-learning parameter was set to zero.  Finally,
-% in the \emph{random-info, att} condition the model was yoked to
-% receive feedback on the same proportion of trials as the contingent
-% model, but the feedback trials were randomly selected and independent
-% of the model's choices. The results of these simulations are plotted
-% in Figure~\ref{fig:model}.
-
-% \begin{figure*}[htp]
-% \centering
-% \includegraphics[width=\textwidth]{figures/modelplot.pdf}
-% \caption{ALCOVE model simulations of approach-avoid decision-making in
-%   five attentional/informational conditions. \emph{Left:} average
-%   score per block of 16 trials; \emph{Center:} proportion of positive
-%   prospects approached; \emph{Right:} proportion of negative prospects
-%   approached. A persistent learning bias develops only when the model
-%   received action-dependent feedback and is endowed with selective
-%   attention. All conditions were simulated for 1000 model runs.}
-% \label{fig:model}
-% \end{figure*}
-
-% As depicted in the left-most panel, agents in the contingent
-% condition  fell into the learning trap and developed a
-% persistent bias, failing to reach perfect performance.  In contrast, 
-% agents in the full-info condition quickly reached
-% peak performance of 12 points per block. The $p(approach|good)$ and
-% $p(approach|bad)$ plots (middle and right panel, respectively) show 
-% that the lower performance of the contingent
-% condition was not due to continued approach of negative, costly
-% prospects, but rather due to the persistent avoidance of some positive
-% prospects.
-
-% When attention learning is removed, the persistent bias disappears and
-% the model reaches near-peak performance by the last block, though it
-% is slower to learn to avoid negative prospects. This should not be
-% interpreted simply as evidence that selective attention is generally
-% harmful in this environment. On the contrary, with full information,
-% selective attention leads to better performance; the full-info,
-% no-att condition performs poorly, overgeneralizing from the negative
-% prospects. It is only when feedback is dependent on current belief and
-% action that selective attention becomes a disadvantage.
-
-% It is also not the case that the contingent condition is biased due
-% to a general lack of information. In the random-info condition, where
-% the model is given the same number of feedback experiences as the
-% contingent condition but distributed evenly over the space of prospects, no
-% bias develops. Thus, the attentional hot stove effect occurs due not to an overall
-% poverty of information but to a specific pattern of behavior
-% which prevents information from being gained about prospects which
-% could correct the model's misallocated attention.
-
-% Finally, it is worth noting that in our simulations we assume that
-% nothing is encoded in the absence of feedback. However, recent
-% experiments have suggested that people might employ constructivist
-% coding in the absence of feedback, essentially storing the exemplar as
-% though the expected outcome had occurred and reinforcing existing
-% beliefs~\citep{henriksson2010coded}. With this
-% coding scheme, we would expect the attentional hot stove effect
-% exhibited by ALCOVE to be even more pronounced.
-
-% \section{Experiment}
-
-% To test the degree to which people are susceptible to the
-% attentional hot stove effect, we performed a simple experiment similar
-% to the category-learning task described above.
-
-% \subsection{Method}
-
-% \subsubsection{Participants.}
-% One hundred one participants were recruited via Amazon Mechanical
-% Turk. Participants received \$1.25 for participation and received a
-% performance-based bonus that ranged up to \$1.80.
-
-% \subsubsection{Stimuli.}
-% Stimuli were computer-generated cartoon bees that varied on four
-% binary dimensions; they had two or six legs, a striped or spotted
-% body, single or double wings, and antennae or no antennae, for a total
-% of 16 unique stimuli (Figure~\ref{fig:stim}). Two of the four
-% dimensions were chosen as relevant, counterbalanced across
-% participants. Of the four possible combinations of values on these two
-% dimensions, one was chosen at random; stimuli with this combination of
-% values were  ``dangerous,'' and the remaining stimuli were
-% ``friendly.''
-
-% \subsubsection{Procedure.}
-% Participants played the role of a beekeeper collecting honey from
-% several beehives. They were told that each hive contained a single
-% variety of bees, and that while most hives contained friendly bees
-% that would give them honey, some
-% hives had been invaded by dangerous bees, which would sting them if
-% they tried to harvest.
-
-% In the learning phase, participants encountered each of the 16 bee
-% varieties 4 times, for a total of 64 trials. They were informed of the
-% number of trials, and a the number of remaining trails was displayed
-% throughout learning. Stimuli were ordered such that
-% every eight stimuli contained two dangerous and six friendly bee
-% varieties, and were otherwise random. On each trial, participants
-% visited a new beehive, and were shown one of the bees in the
-% hive. Based on the bee's appearance, they then had to choose either to
-% attempt to harvest honey from the bee variety in that hive or to avoid
-% the hive. When participants chose to harvest, they received honey and
-% added \$0.02 to their bonus if the bee variety was friendly, but were
-% stung and lost \$0.10 from their bonus if it was dangerous. When
-% participants chose to avoid a hive, they gained \$0.00.  Participants
-% started the game with a bonus of \$0.40.
-
-% Participants were split into two conditions, which differed in 
-% the feedback received upon avoiding a beehive. In the contingent
-% condition, no feedback was provided when a participant avoided
-% a hive. In the full-info condition, participants still gained
-% \$0.00 when they avoided a hive, but were informed of whether the bee
-% variety was friendly or dangerous and of what their payoff would have
-% been had they harvested.
-
-% The learning phase was followed by a surprise test
-% phase.  During the test phase, participants encountered each variety
-% twice and chose to harvest or avoid hives as before, but received no
-% feedback about the outcomes of their actions and were not able to see
-% changes to their bonus. This phase provided a comparison of
-% learning under equivalent conditions.
-
-% After the test phase, participants were informed of their total
-% bonus, and were asked two final questions: ``About what percentage of
-% beehives do you think contained dangerous bees?'' and ``Which features
-% do you think were useful in deciding whether a bee variety was
-% friendly or dangerous?''. For the first question, participants entered
-% a percentage between 0 and 100, and for the
-% second question participants could choose any combination of the four
-% features using check-boxes.
-
-% \begin{figure}[htp]
-% \centering
-% \includegraphics[width=\columnwidth]{figures/stimuli.pdf}
-% \caption{Example stimuli with opposite values on all four binary dimensions.}
-% \label{fig:stim}
-% \end{figure}
-
-% \section{Results and Discussion}
-
-% \subsection{Learning}
-% Learning performance averaged in 16-trial blocks is shown in Figure~\ref{fig:results},
-% \emph{left}, separated by positive and negative prospects. 
-% In the first block of learning, participants in the contingent condition
-% approached all prospects more than those in the full-info condition,
-% $p<.001$.\footnote{All p values are calculated via two-sided permutation test
-%   unless otherwise specified.} This suggests participants valued the information which was
-% gained by approaching in this condition, in line with the results of
-% other recent studies which find that people are information-seeking in
-% simple decision-making tasks~\citep{speekenbrink2014uncertainty,
-%   rich2014value, Wilson2014a}.
-
-% By the last block of learning participants in both conditions rarely approached
-% bad prospects, with no difference between
-% the conditions, $p>.25$. Participants in the full-info condition learned
-% to approach good prospects at a higher rate by the end of
-% learning ($p<.001$), while participants in the contingent condition
-% approached good prospects less frequently ($p=.011$), such that by
-% the final block participants in the full-info condition were
-% significantly more likely to approach a positive prospect than in the
-% contingent condition, $p=.017$.
-
-% \begin{figure*}[htp]
-% \centering
-% \includegraphics[width=\textwidth]{figures/experimentplot.pdf}
-% \caption{\emph{Left:} Proportion of negative and positive prospects
-%   approached by learning block for the full-info and contingent
-%   conditions. Error bars are standard error of the mean. \emph{right:}
-%   Performance in the test phase for participants in each condition,
-%   plotted as proportion of positive prospects approached against
-%   proportion of negative prospects approached and jittered slightly
-%   to increase visibility of overlapping points. The point (1,0) represents
-%   perfect performance, while the black line denotes chance performance.}
-% \label{fig:results}
-% \end{figure*}
-
-% \subsection{Test performance}
-% Performance in the test phase is plotted in
-% Figure~\ref{fig:results}, \emph{right}, which shows each participant's
-% proportion of approaching good and bad prospects.
-% Participants in the full-info condition were significantly more
-% accurate at the test phase, $p=.017$, choosing the correct action on
-% 81.5\% of trials versus 71.5\% for the contingent
-% condition. Interestingly, they did not gain significantly more points
-% on average ($p>.250$), though the difference in median number of
-% points scored approached significance ($p=.106$). This lack of
-% difference in score is due in part to a small subset of participants
-% in the full-info condition who approached all stimuli at a high rate, thus
-% incurring a large cost from the bad prospects.
-
-% The higher accuracy of participants in the full-info condition shows
-% that they better followed the true, 2-feature rule. However, it does
-% not show whether contingent condition participants were less accurate
-% because they followed a uni-dimensional rule, or simply because they
-% were more noisy. To determine the extent to which participants in each
-% condition followed a one-feature rule, we calculated a ``1-feature rule
-% score'' for each participant. This score was determined by calculating
-% the proportion of trials on which each participant followed each of
-% the two relevant 1-feature rules, and then taking the
-% maximum over these two proportions. Participants in the full-info
-% condition had an average 1-feature rule score of $0.74$, while those
-% in the contingent condition had a significantly higher score of
-% $0.83$, $p=.004$. Thus, the
-% difference in accuracy between the full-info and contingent condition
-% participants does not appear to be simply a product of noise due to the
-% latter group's restricted information. Rather, it seems that
-% participants in the contingent condition systematically attended to only one
-% of the two relevant dimensions, thus avoiding a consistent subset of
-% rewarding beehives.
-
-% \subsection{Post-test questions}
-% Participants in the contingent condition responded on average that
-% 35.8\% of prospects were bad, while participants in the full-info
-% condition responded that only 28.2\%, a marginally significant
-% difference ($p=.055$). The true proportion was 25\%. This supports the conjecture that
-% action-dependent feedback can affect a person's beliefs about the
-% environment, and is consistent with the findings of
-% \citet{fazio2004attitude} that approach-avoid learning lead to belief
-% that the environment was more negative than reality. In addition, while 
-% only 22.9\% of participants in
-% the contingent condition identified the right combination of relevant
-% features, 40.4\% of participants in the full-info condition did, a marginally significant difference by Fisher's exact test ($p=.054$).
-
-% In summary, we have provided empirical and computational
-% evidence of an attentional
-% learning trap wherein avoidance behavior promotes the 
-% persistence of false negative beliefs via attentional learning.
-% While the overall effect we report is rather intuitive, it is
-% important to point out that the vast majority of categorization
-% studies have ignored the impact of choice-contingency on
-% learning (essentially focusing exclusive on the ``full info" conditions in our
-% experiment).  In addition, even the traditional, single prospect
-% version of the hot stove effect  has proven
-% surprisingly difficult to produce in laboratory
-% settings~\citep[e.g.,][]{biele2009learning, rich2014value}.
-
-% \section{Using cognitive models to meliorate the hot stove effect}
-
-% Thinking about choice-contingent learning traps in the context of
-% category learning carries several real-world implications. For instance, 
-% screening job applicants usually involves a stream of encounters with many 
-% different prospects that possess a variety of attributes.  The 
-% challenge for a firm is to determine which combinations of attributes tend 
-% to signal a good worker.  As in our experiment, it is clear there is
-% a strong potential for biased hiring rules that screen out many good 
-% candidates due to the selective utilization of certain application features.
-
-% Given these societal concerns, it is interesting to consider if
-% insight from computational models of learning can provide guidance on
-% how best to limit this bias. When the hot stove effect is caused by
-% true stochasticity, it is difficult to prevent
-% \citep{denrell2007adaptive}, though considering the future value of
-% information can limit its severity \citep{rich2014value}. But in cases
-% where the learning trap is primarily attentional, we hypothesize that
-% changes to the environment which disrupt the narrowing of attention
-%  may significantly reduce biases.  In
-% the following section, we explore this issue computationally by asking
-% which manipulations to the learning environment help ALCOVE ``avoid
-% the trap."
-
-% %Second, it provides a potential avenue to ameliorate the effect,
-% %helping people to create more accurate representations of the
-% %enviornment and subsequently make better choices. When the hot stove
-% %effect is caused by true stochasticity in the environment, it is
-% %difficult to prevent. While a proper evaluation of the future value of
-% %gaining information can mitigate the effect in situations where it
-% %would be most costly \citep{rich2014value}, even an
-% %optimal agent exhibits the effect to some degree
-% %\citep{denrell2007adaptive}. But when the hot stove effect is due
-% %instead to learned inattention to relevant dimensions, there is an
-% %increased potential for interventions that could improve long-run
-% %belief and choice. Specifically, changes to the environment that
-% %disrupt the narrowing of attention onto a small number of dimensions
-% %and encourage people to consider a wider set of hypotheses about the
-% %environment could make people less likely to maladaptively avoid positive
-% %prospects. Here, we briefly explore two interventions that could
-% %widen people's attention.
-
-% \subsection{Debiasing interventions}
-
-% In this section, we present two interventions we 
-% hypothesized would mitigate the attentional hot
-% stove effect.  We first describe the intuitions behind the interventions
-% and then provide a modeling analysis testing the effectiveness
-% of each.
-
-% \subsubsection{Individuating prospects.}
-% One possible way to limit the attentional hot stove effect is to make
-% stimuli increasingly distinct and idiosyncratic. When stimuli are more
-% distinctive, people tend to treat them more as individuals and show
-% increased ability to learn identification compared to
-% categorization. While identification learning is more difficult than
-% categorization with generic artificial stimuli
-% \citep{Shepard61learning, love2004sustain}, \citet{Medin1983} found
-% that people were more easily able to pair unique first names than
-% categorical last names with photographs of women's
-% faces. \citet{love2004sustain} found that this phenomenon could be
-% accounted for with the SUSTAIN model of categorization by assuming
-% that the women's faces had many distinctive features beyond those
-% manipulated by the experimenters, which decreased the similarity among
-% stimuli and thus increasing the odds of representing each stimulus
-% individually.
-
-% In an approach-avoid decision-making task, increased individuation of stimuli should
-% make a person less likely to generalize information gained from
-% experience with one prospect to decisions about another. Attention
-% paid to idiosyncratic features will slow the biasing of attention
-% towards a single dimension, giving the person more opportunity to
-% explore a variety of stimuli and learn the true structure of the
-% environment. Essentially, increased individuation of prospects shifts
-% the task away from category-learning, and towards learning about
-% whether to approach individual prospects.  In addition, it may be easier
-% for people to track the history of past rewards with more individually
-% memorable stimuli.
-
-% \subsubsection{Occluding feature information.}
-% A second approach to decreasing the attentional hot stove effect may
-% be, paradoxically, to restrict information by randomly occluding some
-% features of a prospect such that the decision-maker can't
-% observe their values. While this intervention could of course impair a
-% person's decision-making ability, it could actually improve
-% performance in the long run by causing a greater spread of
-% attention. \citet{Taylor2009} found that participants learned more
-% about non-diagnostic features in a category-learning task when
-% features were randomly occluded, and hypothesized that
-% feature-occlusion discourages narrowly selective attention and
-% promotes a broader attentional strategy. In the context of approach
-% decisions, if a person is attending strongly to a dimension which is
-% obscured, he or she may try using other features, which may lead
-% to the discovery that they are relevant. Even when the favored
-% features is not occluded, the possibility of their future absence may
-% cause people to be less quick to look solely at those features
-% \citep{Rehder2005}.
-
-% \begin{figure}[htp]
-% \centering
-% \includegraphics[width=\columnwidth]{figures/interventionplot.pdf}
-% \caption{Average performance of ALCOVE model in a test block after four or
-%   twenty learning blocks, in the contingent and full-info condition and
-%   after two attentional interventions. All conditions simulated for
-%   1000 runs of the model.}
-% \label{fig:interventions}
-% \end{figure}
-
-% \subsection{Modeling debiasing interventions}
-% To test the possible efficacy of these interventions, we performed
-% model simulations comparing them to the contingent and full-info
-% conditions. To modify the model for the individuation condition, we
-% added an extra dimension with 16 nominal values representing
-% idiosyncratic features of each stimulus
-% \citep[following][]{love2004sustain}. For the occluded-dimension condition, on
-% $\frac{1}{4}$ of trials a randomly chosen dimension was masked such
-% that it did not contribute to the model's network activation and its
-% attention weight was not updated.
-
-% The models were simulated for learning phases of two different
-% lengths, 4 blocks and 20 blocks, followed by a one block test
-% phase with no feedback, where no dimensions were masked in the
-% occluded-dimension condition and the individuating dimension was
-% masked in the individuation condition. Performance of the models at
-% test is reported in Figure~\ref{fig:interventions}.
-
-% With only 4 blocks of learning, as in our
-% experiment, the efficacy of the interventions is low. The addition of
-% idiosyncratic features aids learning only slightly, and occluding
-% dimensions actually hurts performance, as the decrease in information
-% from missing dimensions hurts learning more than increasing the spread
-% of attention improves it.
-
-% After a more substantial 20 blocks of learning, the interventions
-% are more effective. Performance in the individuation condition
-% approached that of the full information condition, and performance in
-% the occluded-dimension condition surpasses that of the contingent
-% condition. Future work will aim to evaluate these interventions with
-% human participants, and to develop methods that might speed up their 
-% effectiveness such as maximizing the salience of individuating features.
-
-% \section{Conclusion}
-% Making decisions from experience is an essential part of
-% adaptive cognition, yet such decisions can produce
-% biases in action and belief. Here, we considered one
-% mechanism through which such biases might develop, which we term the
-% attentional hot stove effect.   This effect is a natural consequence of
-% popular category learning models but has so far been largely 
-% ignored.   To the extent that such biases are caused by misallocation 
-% of attention, rather than irreducible stochasticity in the environment, it 
-% may be possible to alter decision-making patterns or the environment itself to
-% facilitate better learning and choice.
+\section{Discussion}
 

### 2017-01-26 analysis/analysis.R

@@ -1,0 +1,174 @@
+library("dplyr")
+library("ggplot2")
+library("tidyr")
+library("lme4")
+
+exploredata <- read.csv('../data/exploration_data_v2_4.csv')
+consumptiondata <- read.csv('../data/consumption_data_v2_4.csv')
+
+questiondata <- read.csv('../data/question_data_v2_4.csv')
+
+questiondata$enjoymentdiff <- questiondata$videoenjoyment - questiondata$sliderenjoyment
+
+mean(questiondata$videoenjoyment)
+mean(questiondata$sliderenjoyment)
+t.test(questiondata$enjoymentdiff)
+
+## exploredata$condition <-  as.factor(exploredata$condition)
+
+exploredata$videoenjoyment <- 0
+exploredata$sliderenjoyment <- 0
+for (id in levels(exploredata$uniqueid)) {
+  exploredata[exploredata$uniqueid == id,"videoenjoyment"] <- questiondata[questiondata$uniqueid == id, "videoenjoyment"]
+  exploredata[exploredata$uniqueid == id,"sliderenjoyment"] <- questiondata[questiondata$uniqueid == id, "sliderenjoyment"]
+  exploredata[exploredata$uniqueid == id,"instructionloops"] <- questiondata[questiondata$uniqueid == id, "instructionloops_instructions.quiz.html"]
+}
+
+exploredata$enjoymentdiff <- (exploredata$videoenjoyment - exploredata$sliderenjoyment) / 10
+
+exploredata <- exploredata %>%
+  mutate(outcome5back=dplyr::lag(outcome, n=5, default=0),
+         outcome1back=dplyr::lag(outcome, n=1, default=0),
+         lastoutcome=ifelse(condition == 1, outcome5back, outcome1back))
+
+practicedata <- exploredata %>% filter(taskType == "practice")
+exploredata <- exploredata %>% filter(taskType != "practice")
+
+exploredata$binnedValue <- floor(exploredata$currentValue * 30)/30
+
+
+exploredata %>% group_by(condition, binnedValue) %>%
+  summarize(response=mean(response)) %>%
+  ggplot(aes(x=binnedValue, y=response, color=as.factor(condition))) +
+  geom_line()
+
+
+ggplot(exploredata, aes(x=currentValue, y=response, group=uniqueid, color=as.factor(condition))) +
+  geom_point(position = position_jitter(width = 0, height = 0.05)) +
[…]

### 2017-01-27 doc/cogsci2017/paper.tex
adding figures, results for Experiment 1, etc

@@
 prevalent, though perhaps in different settings, and that the seemingly general bias
-toward under-exploration is illusory. However, in the current paper we propose
-instead that lab tests of exploratory choice have been missing a key component
+toward under-exploration is illusory. In the current paper, we propose an
+alternative hypothesis: that lab tests of exploratory choice have been missing a key component
 of real-world exploratory choice, namely, the distribution of choices and
@@
 % appears close to optimal \citep{Navarro2016}.
+
+\begin{figure*}
+\centering
+\includegraphics[width=\textwidth]{figures/taskpictures.png}
+\caption{\emph{a}: An example of the slider task. Participants had to move all
+  sliders to ``50'' in 30 seconds. \emph{b}: the video-watching task.
+  Participants had to hold the space bar to watch their chosen video. \emph{c}:
+  The decision-making task. Participants had to choose to run the machine with
+  the current spinner or try a new spinner. If their chosen spinner landed on
+  a gold wedge, they performed the video-watching task instead of the
+  slider task.}
+\label{fig:taskpictures}
+\end{figure*}
 
@@
 in human decision making, and that people are strongly
-biased towards present . For example, people will often prefer a larger, later monetary
-reward to a smaller, sooner reward when both rewards are in the future. But when the
-time until both rewards is reduced so that the sooner reward is immediate, the
-sooner reward seems to loom larger and
-people may switch their preference \citep{Kirby1995}. With non-monetary rewards,
+biased towards  the present. For example, people will often prefer a larger, later monetary
+reward to a smaller, sooner reward when both rewards are in the future, but will
+switch their preference when the time until both rewards is reduced so that the
+sooner reward is immediate \citep{Kirby1995}. With monetary rewards, the delay
+or speed-up required to observe preference reversals is usually several days. With non-monetary rewards,
 such as the cessation of an annoying noise \citep{Solnick1980}, watching a video
@@
 increasing the perceived immediate gains of exploiting relative to the delayed
-gains of exploring. However, to date there has been no experimental research
+gains of exploring. However, there has been no experimental research
 examining how people's bias towards the present affects exploratory choice. In
-part, this is a product of limits in experiment design. In most exploratory
+part, this is by limits in experiment design. In most exploratory
 choice experiments, each choice and outcome takes fewer than 10 seconds to
@@
 and exploratory choice. We develop a novel exploratory choice task using
-real consumption rewards and present two experiments. In the first, we vary
+immediately-consumed rewards and present two experiments. In the first, a pilot
+experiment, we vary
 the expected payoff of exploration across conditions and test whether this affects
-participant choice, to ensure that participants understand the task and respond
+participant choice. The results of this experiment show that participants understand the task and respond
 as predicted to well-studied variables. In the second, we introduce a
-delayed-outcome condition, an intervention only expected to increase exploration
-if people are present biased. While we find no effect of delay on people's
-choices, we discuss plans for future experiments as well as the difficulties of
-studying choice with consumable rewards.
+delayed-outcome condition, which we predicted to increase exploration
+if people are present biased. We find no evidence that delaying outcomes affects
+the rate of exploration. While this may support the conclusion that exploratory
+choice is not susceptible to present bias, we also discuss other possible
+explanations of this null result and plans for future experiments.
 
@@
 
-\begin{figure*}[htp]
+\begin{figure*}
 \centering
@@
   and conditions, six 30-second work periods pass between consecutive choices
-  with a given machine. In Experiment 1 and the immediate condition (left) of
+  with a given machine. \emph{a}: in Experiment 1 and the immediate condition of
   Experiment 2, the outcome for a choice is revealed and occurs immediately
-  after the choice is made. In the delayed condition of Experiment 2 (right),
+  after the choice is made. \emph{b}: in the delayed condition of Experiment 2 (right),
   the outcome of a choice is revealed and occurs after a delay of four work
@@
 The slider task was based of a task previously used by \citet{Gill2012}. In each
-period of the slider task, five horizontal sliders appeared on the screen. Each
+period of the slider task, five horizontal sliders appeared on the screen (Figure~\ref{fig:taskpictures}\emph{a}). Each
 started at a random setting between 0 and 100, with the slider's value
@@
 The video was embedded in the experiment with all user controls (such as
-skipping ahead) disabled. When given access to the video, participants had to keep the
+skipping ahead) disabled (Figure~\ref{fig:taskpictures}\emph{b}). When given access to the video, participants had to keep the
 browser window open and hold down the space bar for the video to play. This allowed us to ensure that
@@
 slider task. After these initial periods, participants gained access to six
-``machines'', each of which had a ``current spinner'' and a ``new spinner''.
+``machines'', each of which had a ``current spinner'' and a ``new spinner'' (Figure~\ref{fig:taskpictures}\emph{c}).
 Before each work period, participants clicked the circled machine,
@@
 made, and affected the next work period, as shown in
-Figure~\ref{fig:machinetimeline} (left). It then ``cooled off'' for the
+Figure~\ref{fig:machinetimeline}\emph{a} and Figure~\ref{fig:machinemappictures}\emph{a}. It then ``cooled off'' for the
 following five periods, as choices were made with the other five machines.
@@
 
+\begin{figure*}
+\centering
+\includegraphics[width=\textwidth]{figures/machinemappictures.png}
+\caption{The machine display seen by participants to allow them to keep track of
+  the value of each of the six machines and the next time it would be ready to
+  make a choice or produce an outcome. \emph{a}: the display seen by
+  participants in Experiment 1 and the immediate condition of Experiment 2.
+  \emph{b}: the display seen by participants in teh delayed condition of
+  Experiment 2.}
+\label{fig:machinemappictures}
+\end{figure*}
+
 \section{Results}
@@
 
-To analyze participants trial-by-trial decision-making behavior, we conducted a hierarchical
+To analyze participants trial-by-trial decision-making, we conducted a hierarchical
 Bayesian logistic regression using the Stan modeling language \citep{stan-software:2015}.
@@
 ready for another choice. The timeline for the delayed condition is shown in
-Figure~\ref{fig:machinetimeline} (right).
+Figure~\ref{fig:machinetimeline}\emph{b}.
 
-\section{results}
+\section{Results}
+
+As in Experiment 1, participants rated the video as more enjoyable () than the
+slider task (), STAT.
+
+To analyze trial-by-trial decision-making, we conducted a hierarchical Bayesian
+logistic regression identical to the one described in the results of Experiment 1.
+
+The results of the regression are shown in Figure~FIGNUM. As in Experiment 1,
+participants are less likely to choose a new spinner when the current spinner
+has a high value, $p>.999$. However, in this experiment we find no effect of
+condition---participants were no more likely to explore a new spinner when there
+was a temporal delay imposed between their choices and the received outcomes.
 

### 2017-01-28 doc/cogsci2017/paper.tex
add abstract and Exp 2 participants and result stats

@@
 
+  Balancing exploration and exploitation is difficult, and under-exploration
+  appears to be a particularly pressing problem. We propose that one possible cause of
+  under-exploration is present bias, in which immediate rewards (like those
+  gained from exploitation) loom larger than future rewards (like those
+  gained from exploration). This possible cause of under-exploration is not
+  addressed by past lab studies, in which choices generally yield token rewards
+  that are converted to money at the end of the experiment, removing the
+  inter-temporal aspect of the decision-making process. In this paper, we
+  develop an exploratory choice task with immediately-consumed rewards. In
+  Experiment 1, we show that people understand the task and respond as expected
+  to standard decision variables. In Experiment 2, we introduce a condition in
+  which there is a temporal delay between choices and outcomes, and test whether
+  this increases exploration by reducing present bias. While we find no effect
+  of the delay, we discuss plans for future experiments to test the validity and
+  robustness of this result.
+
   \textbf{Keywords} time preference, present bias, exploratory choice
@@
 
-The problem of under-exploration seems remarkably common. One of the most severe
+Balancing exploitation of known options with exploration of known options
+presents a difficult cognitive and motivational challenge. The problem of how to
+trade off between these two conflicting goals has created whole fields of study
+\citep{Mehlhorn2015, sutton1998reinforcement}, and in many cases achieving an
+optimal balance is computationally intractible \citep{Guez2013}. Given these
+facts, one might expect people to often deviate from optimal exploratory choice
+in both directions, perhaps exploring too much in some situations but too little
+in others.
+
+Notably, though, while over-exploration does occur, a bias towards under-exploration in particular seems to be
+observed across a wide range of domains. One of the most severe
 expressions of under-exploration is learned helplessness, in which an organism
@@
 environment is uncontrollable, and thus ceases to take actions that might allow
-it to regain its sense of control. Learned helplessness has been proposed to
+it to discover that it can in fact exert control. Learned helplessness has been proposed to
 underly most notably depression \citep{Abramson1978, Abramson1989} but also
 problems ranging from difficulties in school CITE to poverty CITE.
-
 While the cognitive appraisal of experienced events affects the development of
@@
 account for the tendency to under-explore.
-
-% In ``observe or bet'' tasks in which people choose to either
-% obtain information or reward on each trial, people appear to collect much more
-% information than is needed \citep{tversky1966information}. But, in settings
-% where the underlying reward probabilities might change, information-seeking
-% appears close to optimal \citep{Navarro2016}.
 
@@
 
+\begin{figure*}
+\centering
+\includegraphics[width=\textwidth]{figures/machinemappictures.png}
+\caption{The machine display seen by participants to allow them to keep track of
+  the value of each of the six machines and the next time it would be ready to
+  make a choice or produce an outcome. \emph{a}: the display seen by
+  participants in Experiment 1 and the immediate condition of Experiment 2.
+  \emph{b}: the display seen by participants in teh delayed condition of
+  Experiment 2.}
+\label{fig:machinemappictures}
+\end{figure*}
+
 In the main part of the experiment, participants completed a total of 70 work
@@
 
-\begin{figure*}
-\centering
-\includegraphics[width=\textwidth]{figures/machinemappictures.png}
-\caption{The machine display seen by participants to allow them to keep track of
-  the value of each of the six machines and the next time it would be ready to
-  make a choice or produce an outcome. \emph{a}: the display seen by
-  participants in Experiment 1 and the immediate condition of Experiment 2.
-  \emph{b}: the display seen by participants in teh delayed condition of
-  Experiment 2.}
-\label{fig:machinemappictures}
-\end{figure*}
-
 \section{Results}
@@
 
+Forty participants completed the experiment, which was conducted over
+Amazon Mechanical Turk using the psiTurk framework \citep{Gureckis2015a}.
+Participants were paid \$5.00 for their participation, with a performance-based
+bonus of up to \$3.00. All participants received the full \$3.00 bonus.
+
 \subsubsection{Design and procedure.}
@@
 
-As in Experiment 1, participants rated the video as more enjoyable () than the
-slider task (), STAT.
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{figures/exp2results.pdf}
+\caption{Model-based estimates of participants' probability of choosing a new
+  spinner for different values of the current spinner in Experiment 2. Thick
+  lines and shaded regions indicate the mean and 95\% posterior interval for the
+  population-level parameters, while the thin lines indicate the mean posterior
+  parameters for each of the 40 individual participants. Participants in the
+  delayed-outcome condition were no more
+  likely to explore at a given current-spinner value than those in the
+  immediate-outcome condition.}
+\label{fig:exp2results}
+\end{figure}
+
+As in Experiment 1, participants rated the video as more enjoyable on average (5.65 out of 7) than the
+slider task (3.13 out of 7), $t(39)=9.26$, $p<.001$.
 
@@
 
-The results of the regression are shown in Figure~FIGNUM. As in Experiment 1,
-participants are less likely to choose a new spinner when the current spinner
-has a high value, $p>.999$. However, in this experiment we find no effect of
-condition---participants were no more likely to explore a new spinner when there
+The results of the regression are shown in Figure~\ref{fig:exp2results}. As in Experiment 1,
+participants were less likely to choose a new spinner when the current spinner
+has a high value, $p>.999$. However, in this experiment we found no evidence of
+an effect of condition, with a posterior probability of $p=.558$ that the effect of condition was
+greater than 0. This means that participants were no more likely to explore a new spinner when there
 was a temporal delay imposed between their choices and the received outcomes.
+However, there may be a small interaction between current spinner value and
+condition, such that participants in the delayed condition were less sensitive
+to the value of the spinner when making their choices. This might indicate that
+the delayed condition was confusing to some participants, as a few individuals
+(as seen in Figure~\ref{fig:exp2results}) changed their behavior very little
+across current-spinner values. The posterior probability that this interaction
+was above 0 is $p=.934$.
 
 \section{Discussion}
+
+In this paper we presented a novel task using immediately-consumed rewards that
+may be used to study present bias in the context of exploratory choice. In
+Experiment 1 we showed that participants appeared to understand the task and
+respond reasonably to maximize their time in the video-watching task. They are
+more likely to explore when the current exploitable option (current spinner) has
+low value, and change their degree of exploration based on the range of possible
+outcomes from choosing a new spinner. However, in Experiment 2 we found that
+participants did not explore any more when there was a temporal delay between
+choices and outcomes, even though we expected this delay to reduce the present
+bias towards the greater immediate rewards of exploitation.
+
+One explanation for our results is that people are in fact less affected by
+present bias in the realm of exploratory choice than they are in other
+inter-temporal choices.
 

### 2017-01-29 doc/cogsci2017/paper.tex
paper first draft - finished discussion

@@
 Under-exploration also seems to occur in the development of complex skills, such
-as flying a plane or playing a sport CITE. In these settings, an ``emphasis
+as flying a plane or playing a sport \citep{Gopher1989}. In these settings, an ``emphasis
 change'' training method in which the focus of practice is repeatedly changed
@@
 continually explore the performance space to find more effective cognitive and
-motor routines CITE. Without these interventions, people often enter a ``local
-maximum'' in which exploration decreases and performance plateaus CITE.
+motor routines. Without these interventions, people often enter a ``local
+maximum'' in which exploration decreases and performance plateaus \citep{Yechiam2001}.
 
@@
 Thirty-one participants completed the experiment, which was conducted over
-Amazon Mechanical Turk using the psiTurk framework \citep{Gureckis2015a}.
+Amazon Mechanical Turk (AMT) using the psiTurk framework \citep{Gureckis2015a}.
 Participants were paid \$5.00 for their participation, with a performance-based
@@
 Forty participants completed the experiment, which was conducted over
-Amazon Mechanical Turk using the psiTurk framework \citep{Gureckis2015a}.
+AMT using the psiTurk framework \citep{Gureckis2015a}.
 Participants were paid \$5.00 for their participation, with a performance-based
@@
 present bias in the realm of exploratory choice than they are in other
-inter-temporal choices.
+inter-temporal choices. This might be the case if exploratory choice is driven
+less by an explicit present--future tradeoff and more by intrinsic ``exploration
+bonuses'' \citep{kakade2002dopamine} or motivational states such as curiosity
+\citep{Berlyne1966}. If exploration is driven by these mechanisms, then the
+reward of exploration might be exploration itself rather than its later
+consequences, breaking the intertemporal tradeoff between exploration and
+exploitation.
+
+However, given the early stage of the present work, we are not ready to make
+strong claims about the role of present bias in exploratory choice. First, while
+participants enjoyed the video more than the slider task, we need to perform a
+follow-up study to ensure that this difference in enjoyment was strong enough to
+create a bias towards immediate video-watching in the time scale used in our
+experiments. To test this, we might use a simple ``concurrent schedule''
+experiment in which two buttons produce video-watching time at for different
+lengths (or with different probabilities) and at different delays. In this
+setting, we would predict a bias towards the button that produced earlier
+video-watching time, even if the video time was shorter or less frequent.
+\citet{Navarick1998} used a similar paradigm to demonstrate present bias in
+watching cartoons, though the bias might be reduced in participants on AMT due
+to the greater availability of environmental distractors during non-video time.
+
+We also may need to improve our decision-making task. As discussed, some
+participants may have found the delayed condition confusing, and this problem
+might be fixed through clearer instructions or interface. Additionally, the
+future encounters with each machine may not have been sufficiently salient,
+reducing the perceived need for exploration in both conditions.
+
+While introduces difficulties not found in standard decision-making tasks, we
+remain excited about the possibility of creating a paradigm for the use of
+immediately-consumable rewards in dynamic decision-making tasks. Such a paradigm
+could allow not just a test of whether present bias affects exploratory choice,
+but a greater focus on a variety of motivational factors often neglected by
+cognitive science.
 

### 2017-01-29 analysis/analysis.R
switching analyses back to immediate-delay exp

@@ -3,10 +3,10 @@ library("ggplot2")
 library("tidyr")
 library("lme4")
 
-exploredata <- read.csv('../data/exploration_data_v2_4.csv')
-consumptiondata <- read.csv('../data/consumption_data_v2_4.csv')
+exploredata <- read.csv('../data/exploration_data_v2_3.csv')
+consumptiondata <- read.csv('../data/consumption_data_v2_3.csv')
 
-questiondata <- read.csv('../data/question_data_v2_4.csv')
+questiondata <- read.csv('../data/question_data_v2_3.csv')
 
 questiondata$enjoymentdiff <- questiondata$videoenjoyment - questiondata$sliderenjoyment
 
@@ -155,14 +155,15 @@ individuals_w_predictions$condition <- factor(individuals_w_predictions$conditio
 ggplot() + geom_line(data=individuals_w_predictions, aes(x=x, y=y, group=subject, color=condition), size=.5, alpha=.5) +
   geom_line(data=params_w_predictions, aes(x=x, y=y, color=condition), size=2) +
   geom_ribbon(data=params_w_predictions, aes(x=x, y=y, ymin=lower, ymax=upper, group=condition), alpha=.3) +
-  scale_color_discrete(name="condition", labels=c("low-explore", "high-explore")) +
+  scale_color_discrete(name="condition", labels=c("immediate", "delayed")) +
   scale_x_continuous(breaks=c(0.4, 0.6, 0.8, 1), labels=c("40%", "60%", "80%", "100%")) +
   xlab("current-spinner value") +
   ylab("p(choose new spinner)") +
   theme_minimal() +
   coord_cartesian(xlim = c(.333, 1))+
   theme(legend.position="bottom")
-ggsave(filename="../doc/cogsci2017/figures/exp1results.pdf", width=5, height=3.5, useDingbats=F)
+
+ggsave(filename="../doc/cogsci2017/figures/exp2results.pdf", width=5, height=3.5, useDingbats=F)
 
 
 ## fit <- stan(file="stan_models/regression_lastoutcome.stan", data=standata, iter=1000)

### 2017-01-29 analysis/get_data.R

@@ -17,7 +17,7 @@ data = fetch(rs, n=-1)
 
 ## data <- data[data$status %in% c(4, 7),]
 data <- data[data$status %in% c(3, 4, 7),]
-data <- data[data$codeversion %in% c("2.4"),]
+data <- data[data$codeversion %in% c("2.3"),]
 
 datastrings <- data$datastring
 datastrings_json <- sapply(datastrings, fromJSON, simplify=F)
@@ -62,8 +62,8 @@ consumptiondata <- Filter(function(x)!all(is.na(x)), consumptiondata)
 exploredata <- Filter(function(x)!all(is.na(x)), exploredata)
 
 
-write.csv(exploredata, file="../data/exploration_data_v2_4.csv")
-write.csv(consumptiondata, file="../data/consumption_data_v2_4.csv")
+write.csv(exploredata, file="../data/exploration_data_v2_3.csv")
+write.csv(consumptiondata, file="../data/consumption_data_v2_3.csv")
 
 
 questiondata <- sapply(datastrings_json,
@@ -79,4 +79,4 @@ names(questiondata) <- NULL
 questiondata <- as.data.frame(do.call("rbind.data.frame", questiondata))
 rownames(questiondata) <- NULL
 
-write.csv(questiondata, file="../data/question_data_v2_4.csv")
+write.csv(questiondata, file="../data/question_data_v2_3.csv")

### 2017-01-29 analysis/models.R

@@ -77,22 +77,7 @@ optimal_values_indefinite <- function (discount, minreward=4, maxreward=12, pbad
        )
 }
 
-test <- optimal_values_indefinite(.8, minreward=4, maxreward=12, pbad=.33, badcost=0)
-
-df <- data.frame(reward_exploit=test$reward_exploit)
-df$v_diff <- test$reward_diff + test$VoI_diff
-df$v_diff_9 <- test$reward_diff + .9 * test$VoI_diff
-df$v_diff_7 <- test$reward_diff + .7 * test$VoI_diff
-df$v_diff_5 <- test$reward_diff + .5 * test$VoI_diff
-df$v_diff_0 <- test$reward_diff
-
-df <- df %>% gather(bias, diff, -reward_exploit) %>%
-  filter(reward_exploit > 0)
-
-ggplot(df, aes(x=reward_exploit, y=diff, group=bias)) + geom_line(aes(color=bias))
-
-
-test <- optimal_values(5/6, 10)
+test <- optimal_values(5/6, 10, minreward=25, maxreward=100, pbad=0, badcost=0)
 df <- melt(test$exploit_values)
 df <- setNames(df, c("trial", "rewardIdx", "exploit_value"))
 df$explore_value <- melt(test$explore_values)[,3]
@@ -109,7 +94,8 @@ df$v_diff_0 <- df$reward_diff
 df <- df %>%
   select(-reward_diff, -VoI_diff, -exploit_value, -explore_value) %>%
   gather(bias, diff, -trial, -rewardIdx) %>%
-  filter(rewardIdx > 1)
+  filter(rewardIdx > 1) %>%
+  filter(trial == 1)
 
 ## plot of how optimal effect changes as the end of the experiment is reached
-ggplot(df, aes(x=rewardIdx, y=diff, group=bias)) + geom_hline(yintercept=0) + geom_line(aes(color=bias)) + ylim(-7, 5)+ facet_wrap(~trial)
+ggplot(df, aes(x=rewardIdx, y=diff, group=bias)) + geom_hline(yintercept=0) + geom_line(aes(color=bias)) + xlim(0, 100) + ylim(-60, 60)+ facet_wrap(~trial)

### 2017-01-30 doc/cogsci2017/paper.tex
add missing citations on learned helplessness

@@
 underly most notably depression \citep{Abramson1978, Abramson1989} but also
-problems ranging from difficulties in school CITE to poverty CITE.
+problems ranging from difficulties in school \citep{Diener1978} to poverty \citep{Evans2005}.
 While the cognitive appraisal of experienced events affects the development of

### 2017-01-31 doc/cogsci2017/paper.tex
some comments on a first quick read…

@@
 
-\title{Present bias in exploratory choice}
+\title{Does a present bias influence exploratory choice?}
 
@@
 
-  Balancing exploration and exploitation is difficult, and under-exploration
-  appears to be a particularly pressing problem. We propose that one possible cause of
-  under-exploration is present bias, in which immediate rewards (like those
+  Balancing exploration and exploitation is difficult, and across a wide
+  variety of situations under-exploration of new alternatives appears prevalent.
+  We propose that one possible cause of
+  under-exploration is present bias, whereby immediate rewards (like those
   gained from exploitation) loom larger than future rewards (like those
   gained from exploration). This possible cause of under-exploration is not
-  addressed by past lab studies, in which choices generally yield token rewards
+  addressed by past studies, in which choices generally yield token rewards
   that are converted to money at the end of the experiment, removing the
@@
   this increases exploration by reducing present bias. While we find no effect
-  of the delay, we discuss plans for future experiments to test the validity and
-  robustness of this result.
+  of the delay, \hl{we discuss plans for future experiments to test the validity and
+  robustness of this result.}
 
@@
 those with a more cognitive orientation \citep{Jacobson1996}.
+
+\hl{This is good but kind of goes deep into these topics which aren't that central
+to the current paper... i.e., learned helplessness per-se.  it can be a little distracting
+and adds a lot of citations.  incidently the underexploration of product categories
+seems relevant as well}
 
@@
 Under-exploration is often proposed to be the result of myopia: the
-underweighting of temporally distant rewards relative to close ones \citep{March1991, Levinthal1993}. While
+underweighting of temporally distant rewards relative to close ones \citep{March1991, Levinthal1993}. 
+\hl{does this someone undermine the novelty of your framing/prediction?}
+While
 choosing actions that are expected to immediately maximize reward is inherently
@@
 is aware of the future but simply doesn't care as much about it.
-A wealth have studies have shown that this form of myopia is an important factor
+A wealth of studies have shown that this form of myopia is an important factor
 in human decision making, and that people are strongly
@@
 examining how people's bias towards the present affects exploratory choice. In
-part, this is by limits in experiment design. In most exploratory
+part, this is due to limits in experiment design. In most exploratory
 choice experiments, each choice and outcome takes fewer than 10 seconds to
@@
 
-In the present study, we describe our ongoing efforts to address the lack of research on present bias
-and exploratory choice. We develop a novel exploratory choice task using
-immediately-consumed rewards and present two experiments. In the first, a pilot
-experiment, we vary
+In the present study, we describe our ongoing efforts to study the effect of present bias
+on exploratory choice. We develop a novel exploratory choice task using
+immediately-consumed rewards and present two experiments. In the first, we vary
 the expected payoff of exploration across conditions and test whether this affects
@@
 delayed-outcome condition, which we predicted to increase exploration
-if people are present biased. We find no evidence that delaying outcomes affects
+if people are present biased. However, we find no strong evidence that delaying outcomes affects
 the rate of exploration. While this may support the conclusion that exploratory
@@
 bonus of up to \$3.00. All participants received the full \$3.00 bonus.
+\hl{Maybe mention randomization proceedure to the two conditions?}
 
@@
 \end{figure*}
+
+\hl{Complex task to follow...  it might make sense to overview it and then
+divide into subsections.  like "consuption tasks" "choice tasks", etc...}
 
@@
   \emph{b}: the display seen by participants in teh delayed condition of
-  Experiment 2.}
+  Experiment 2. \hl{I wonder if you draw and arrow to show the movement of the machines in the queue?}}
 \label{fig:machinemappictures}
@@
 ``machines'', each of which had a ``current spinner'' and a ``new spinner'' (Figure~\ref{fig:taskpictures}\emph{c}).
-Before each work period, participants clicked the circled machine,
+Before each work period, participants clicked the circled machine, \hl{Unclear what circled machine means}
 were shown the machine and made a choice about which spinner to run. Based on their choices, there was a
@@
 participant to watch their chosen video instead.
+\hl{I wonder if you just say more directly that you spin either an existing dial (exploit) or a new
+dial (explore).  If the machine came up yellow it would do the "work" for you and you could watch
+videos.  If not they you had to do the slider tasks yourself}
 
@@
 spinner, the new spinner was ``saved'' and the current spinner was
-updated to the new spinner.  This meant that
+updated to the new spinner.  \emph{This meant that
 while choosing a new spinner carried risk, it could carry long-term
-benefits as the current spinner could be improved from its initial value.
+benefits as the current spinner could be improved from its initial value.}
 Importantly, the value of choosing a new spinner was higher in the high-spinner
@@
 addition, before beginning the full experiment, participants completed two
-practice phase. First, they performed several trials of practice using the
+practice phases. First, they performed several trials of practice using the
 machines, with the actual work periods removed. Then, they performed three work
@@
 The results of the regression are shown in Figure~\ref{fig:exp1results}.
-Because we observed few trails for each participant, and the current-spinner
+Because we observed few trials for each participant, and the current-spinner
 values observed for each participant depend on chance and the participant's
@@
 
+\hl{A little discussion here might help frame things}
+
 \section{Experiment 2}
@@
 Forty participants completed the experiment, which was conducted over
-AMT using the psiTurk framework \citep{Gureckis2015a}.
+AMT using the psiTurk framework.
 Participants were paid \$5.00 for their participation, with a performance-based
 bonus of up to \$3.00. All participants received the full \$3.00 bonus.
+\hl{Randomization to conditions}
 
@@
 Experiment 2 followed the design of Experiment 1, with the following changes. In
-both conditions, the possible winning proportion of a new spinner ranged fro 0\% to 100\% of
+both conditions, the possible winning proportion of a new spinner ranged from 0\% to 100\% of
 the spinner. In the immediate condition the machine ran immediately after the
@@
 
-While introduces difficulties not found in standard decision-making tasks, we
-remain excited about the possibility of creating a paradigm for the use of
-immediately-consumable rewards in dynamic decision-making tasks. Such a paradigm
-could allow not just a test of whether present bias affects exploratory choice,
-but a greater focus on a variety of motivational factors often neglected by
-cognitive science.
+%%% this is a little too apologetic. who cares about our excitement! ;)
+%While introduces difficulties not found in standard decision-making tasks, we
+%remain excited about the possibility of creating a paradigm for the use of
+%immediately-consumable rewards in dynamic decision-making tasks. Such a paradigm
+%could allow not just a test of whether present bias affects exploratory choice,
+%but a greater focus on a variety of motivational factors often neglected by
+%cognitive science.
 

### 2017-02-01 doc/cogsci2017/paper.tex
making cogsci paper edits suggested by Todd

@@
 Notably, though, while over-exploration does occur, a bias towards under-exploration in particular seems to be
-observed across a wide range of domains. One of the most severe
-expressions of under-exploration is learned helplessness, in which an organism
+observed across a wide range of domains. In learned helplessness, for example, an organism
 experiences the absence of contol over the environment, learns that the
 environment is uncontrollable, and thus ceases to take actions that might allow
-it to discover that it can in fact exert control. Learned helplessness has been proposed to
-underly most notably depression \citep{Abramson1978, Abramson1989} but also
-problems ranging from difficulties in school \citep{Diener1978} to poverty \citep{Evans2005}.
+it to discover that it can in fact exert control. % Learned helplessness has been proposed to
+% underly most notably depression \citep{Abramson1978, Abramson1989} but also
+% problems ranging from difficulties in school \citep{Diener1978} to poverty \citep{Evans2005}.
 While the cognitive appraisal of experienced events affects the development of
 learned helplessness \citep{Abramson1978}, patterns of exploration clearly play
-a role as well \citep{Huys2009, Teodorescu2014a}. In the case of depression,
-interventions aimed at increasing the exploration of activities this might
-produce a sense of pleasure and mastery have been found to be as effective as
-those with a more cognitive orientation \citep{Jacobson1996}.
-
-\hl{This is good but kind of goes deep into these topics which aren't that central
-to the current paper... i.e., learned helplessness per-se.  it can be a little distracting
-and adds a lot of citations.  incidently the underexploration of product categories
-seems relevant as well}
+a role as well \citep{Huys2009, Teodorescu2014a}. % In the case of depression,
+% interventions aimed at increasing the exploration of activities this might
+% produce a sense of pleasure and mastery have been found to be as effective as
+% those with a more cognitive orientation \citep{Jacobson1996}.
 
@@
 
-There are many other areas in which under-exploration is less clearly
+In many other areas under-exploration is less clearly
 established, but is suspected to play a role in maladaptive behavior.
@@
 
-Under-exploration is often proposed to be the result of myopia: the
+Past theoretical work has proposed that Under-exploration may be the result of myopia: the
 underweighting of temporally distant rewards relative to close ones \citep{March1991, Levinthal1993}. 
-\hl{does this someone undermine the novelty of your framing/prediction?}
+% \hl{does this someone undermine the novelty of your framing/prediction?}
 While
@@
 
-The the second form of myopia is that the decision-maker
+The second form of myopia is that the decision-maker
 is aware of the future but simply doesn't care as much about it.
@@
 
+In our experimental paradigm, participants made a series of explore--exploit
+decisions, where each decision determined whether the following 30-second
+consumption task would be unpleasant (moving sliders) or pleasant (watching a
+Youtube video). In Experiment 1, we sought to confirm that participants prefered
+the video task to the slider task and would seek to increase video-watching
+time, and that people were sensitive to explicit, exogenous changes to the value
+of exploration. The experiment consisted of two conditions---in the high-explore
+condition, the average payoff of exploration was high, and in the low-explore
+condition, it was low. We predicted that participants in the high-explore condition
+would make more exploratory choices.
+
 \subsection{Method}
@@
 bonus of up to \$3.00. All participants received the full \$3.00 bonus.
-\hl{Maybe mention randomization proceedure to the two conditions?}
+Participants were pseudo-randomly counterbalanced across two conditions.
 
@@
 
-\hl{Complex task to follow...  it might make sense to overview it and then
-divide into subsections.  like "consuption tasks" "choice tasks", etc...}
+% \hl{Complex task to follow...  it might make sense to overview it and then
+% divide into subsections.  like "consuption tasks" "choice tasks", etc...}
 
+\emph{Consumption tasks.}
 Participants were informed that their job was to perform a monotonous slider task that
@@
 The slider task was based of a task previously used by \citet{Gill2012}. In each
-period of the slider task, five horizontal sliders appeared on the screen (Figure~\ref{fig:taskpictures}\emph{a}). Each
+period of the slider task, five horizontal sliders appeared on the screen (Figure~\ref{fig:taskpictures}a). Each
 started at a random setting between 0 and 100, with the slider's value
@@
 The video was embedded in the experiment with all user controls (such as
-skipping ahead) disabled (Figure~\ref{fig:taskpictures}\emph{b}). When given access to the video, participants had to keep the
+skipping ahead) disabled (Figure~\ref{fig:taskpictures}b). When given access to the video, participants had to keep the
 browser window open and hold down the space bar for the video to play. This allowed us to ensure that
@@
   the value of each of the six machines and the next time it would be ready to
-  make a choice or produce an outcome. \emph{a}: the display seen by
-  participants in Experiment 1 and the immediate condition of Experiment 2.
-  \emph{b}: the display seen by participants in teh delayed condition of
-  Experiment 2. \hl{I wonder if you draw and arrow to show the movement of the machines in the queue?}}
+  make a choice or produce an outcome. Gray arrows have been added to depict the
+  counterclockwise movement of machines around the display after each work
+  period.
+  \emph{a}: the display seen by participants in Experiment 1 and the immediate condition of Experiment 2.
+  \emph{b}: the display seen by participants in the delayed condition of
+  Experiment 2.}
 \label{fig:machinemappictures}
@@
 
-In the main part of the experiment, participants completed a total of 70 work
+Participants completed a total of 70 work
 periods. For the first 10 work periods, participants simply clicked a button to begin the
 slider task. After these initial periods, participants gained access to six
-``machines'', each of which had a ``current spinner'' and a ``new spinner'' (Figure~\ref{fig:taskpictures}\emph{c}).
-Before each work period, participants clicked the circled machine, \hl{Unclear what circled machine means}
-were shown the machine and made a choice about which spinner to run. Based on their choices, there was a
-chance the machine would perform the work task for the participant, allowing the
-participant to watch their chosen video instead.
-\hl{I wonder if you just say more directly that you spin either an existing dial (exploit) or a new
-dial (explore).  If the machine came up yellow it would do the "work" for you and you could watch
-videos.  If not they you had to do the slider tasks yourself}
+machines that could potentially complete the slider task for the participant,
+allowing the participant to watch their chosen video instead. However, the
+machines did not always function, and participants had to make a decision about
+how to set the machine before each use.
+% , each of which had a ``current spinner'' (exploit) and a ``new
+% spinner'' (explore; Figure~\ref{fig:taskpictures}\emph{c}).
+% Before each work period, participants
+% were shown a machine and made a choice about which spinner to run. The spinner
+% then spun, as described below, and if it landed on a ``gold'' wedge then the machine performed the work task for the participant,
+%  allowing the participant to watch their chosen video instead.
 
-The machine ran immediately after the choice was
-made, and affected the next work period, as shown in
-Figure~\ref{fig:machinetimeline}\emph{a} and Figure~\ref{fig:machinemappictures}\emph{a}. It then ``cooled off'' for the
-following five periods, as choices were made with the other five machines.
-
-At each choice, participants had to select between two options: run the
-machine's ``current spinner'' or run a ``new spinner''. Each machine had two
-circular spinners with arrows at the top. The current spinner was split into
+\emph{Decision-making task.}
+Following the initial 10 periods, participants were shown a machine before each work period and had to select
+between two circular spinners with arrows at the top, as shown in Figure~\ref{fig:taskpictures}c: the ``current spinner''
+(exploit) and the ``new spinner'' (explore).
+The current spinner was split into
 five black and five gold wedges. If a participant chose the
 current spinner, it spun (immediately or after four periods, depending
-on the condition) and, if the arrow landed on the gold wedge, the machine
+on the condition) and, if the arrow landed on gold, the machine
 worked and the participant could watch the video. Initially, the
@@
 The new spinner initially showed a question mark. If a participant chose the
-new spinner, then (immediately or after four periods) a new spinner was
+new spinner, then a new spinner was
 created and appeared on the machine. In the high-explore condition, the new spinner's gold wedges could cover
@@
 anywhere from 0\% of the spinner to 75\% of the spinner. The new spinner then
-spun and, if the arrow landed on a gold wedge, the machine worked.
-Additionally, if the new spinner had a greater gold area than the current
+spun and, if the arrow landed on gold, the machine worked.
+
+Critically, if the new spinner had a greater gold area than the current
 spinner, the new spinner was ``saved'' and the current spinner was
-updated to the new spinner.  \emph{This meant that
-while choosing a new spinner carried risk, it could carry long-term
-benefits as the current spinner could be improved from its initial value.}
-Importantly, the value of choosing a new spinner was higher in the high-spinner
-condition, because the range of outcomes was shifted towards higher values.
+updated to the new spinner.  This created an explore--exploit tradeoff in which 
+choosing a new spinner carried immediate risk, but could carry long-term
+benefits by improving the current spinner from its initial value.
+% Importantly, the value of choosing a new spinner was higher in the high-spinner
+% condition, because the range of outcomes was shifted towards higher values.
+
+The machine ran immediately after the choice was
+made, and affected the next work period, as shown in
+Figure~\ref{fig:machinetimeline}a and Figure~\ref{fig:machinemappictures}a. It then ``cooled off'' for the
+following five periods, as choices were made with the other five machines.
 
@@
 
-During the machine choices, participants had access to an ``info'' button at the
-bottom of the screen that provided reminders about the dynamics of the task. In
-addition, before beginning the full experiment, participants completed two
+\emph{Training, incentives, and post-experiment questions.}
+Before beginning the full experiment, participants completed two
 practice phases. First, they performed several trials of practice using the
@@
 periods practicing the slider task.
+During the machine choices, participants had access to an ``info'' button at the
+bottom of the screen that provided reminders about the dynamics of the task.
 
@@
 
-\hl{A little discussion here might help frame things}
+These results show that participants' behavior in this task is
+similar to behavior in other exploratory choice tasks. Participants explored more
+when the current exploitative option is poor, as well as when the possible
+outcomes of exploration are good. While this experiment did not address whether the
+consumption rewards produced a bias towards the present (a topic returned to in the
+discussion), it did showed that participants enjoyed the video more than the slider
+task, and that they made choices to obtain more video time. In the following
+experiment, we introduce a delayed-outcome version of the task and predict that, if the
+consumption rewards induce a present bias, this version will increase exploration.
 
@@
 bonus of up to \$3.00. All participants received the full \$3.00 bonus.
-\hl{Randomization to conditions}
+Participants were pseudo-randomly counterbalanced across two conditions.
 
@@
 both conditions, the possible winning proportion of a new spinner ranged from 0\% to 100\% of
-the spinner. In the immediate condition the machine ran immediately after the
+the spinner. In the immediate condition, the machine ran immediately after the
 choice was made, as in Experiment 1. In the delayed condition, each machine was
@@
 ``process'' for four work periods, thus delaying
-the outcome by over 2 minutes. The participant then clicked the machine again to
-return to it and observe its outcome, and either perform the slider task or
+the outcome by over 2 minutes Figure~\ref{fig:machinemappictures}b. The participant then returned to the
+machine to observe its outcome and either perform the slider task or
 watch the video. The machine then cooled of for a single period before being
 ready for another choice. The timeline for the delayed condition is shown in
-Figure~\ref{fig:machinetimeline}\emph{b}.
+Figure~\ref{fig:machinetimeline}b.
 
@@
 \textbf{Acknowledgments} 
+\hl{FUNDING?}
 \end{small}

### 2017-02-01 doc/cogsci2017/paper.tex
change organization so figure doesn't take up whole column

@@
 
-Past theoretical work has proposed that Under-exploration may be the result of myopia: the
+Past theoretical work has proposed that under-exploration may be the result of myopia: the
 underweighting of temporally distant rewards relative to close ones \citep{March1991, Levinthal1993}. 
@@
 
-\begin{figure}
-\centering
-\includegraphics[width=\columnwidth]{figures/exp1results.pdf}
-\caption{Model-based estimates of participants' probability of choosing a new
-  spinner for different values of the current spinner in Experiment 1. Thick
-  lines and shaded regions indicate the mean and 95\% posterior interval for the
-  population-level parameters, while the thin lines indicate the mean posterior
-  parameters for each of the 31 individual participants. Participants in the
-  high-explore condition were more
-  likely to explore at a given current-spinner value than those in the
-  low-explore condition.}
-\label{fig:exp1results}
-\end{figure}
-
 Participants rated their enjoyment of the video at 5.77 out of 7, on average,
@@
 effect of current-spinner value.
+
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{figures/exp1results.pdf}
+\caption{Model-based estimates of participants' probability of choosing a new
+  spinner for different values of the current spinner in Experiment 1. Thick
+  lines and shaded regions indicate the mean and 95\% posterior interval for the
+  population-level parameters, while the thin lines indicate the mean posterior
+  parameters for each of the 31 individual participants. Participants in the
+  high-explore condition were more
+  likely to explore at a given current-spinner value than those in the
+  low-explore condition.}
+\label{fig:exp1results}
+\end{figure}
 

### 2017-02-24 doc/cogsci2017/paper.tex
added a few items
edits to the cogsci paper I forgot to commit

@@
   Balancing exploration and exploitation is difficult, and across a wide
-  variety of situations under-exploration of new alternatives appears prevalent.
+  variety of situations under-exploration of uncertain alternatives appears prevalent.
   We propose that one possible cause of
@@
   which there is a temporal delay between choices and outcomes, and test whether
-  this increases exploration by reducing present bias. While we find no effect
-  of the delay, \hl{we discuss plans for future experiments to test the validity and
-  robustness of this result.}
+  this increases exploration by reducing present bias. We find no evidence of an
+  effect of delay on exploration, and discuss possible interpretations of this result.
 
@@
 
-Balancing exploitation of known options with exploration of known options
+Balancing exploitation of known options with exploration of uncertain options
 presents a difficult cognitive and motivational challenge. The problem of how to
@@
 \citep{Mehlhorn2015, sutton1998reinforcement}, and in many cases achieving an
-optimal balance is computationally intractible \citep{Guez2013}. Given these
+optimal balance is computationally intractable \citep{Guez2013}. Given these
 facts, one might expect people to often deviate from optimal exploratory choice
 in both directions, perhaps exploring too much in some situations but too little
-in others.
+in others. Notably, though, while over-exploration does occur, a bias towards under-exploration in particular seems to be
+observed across a wide range of domains.
 
-Notably, though, while over-exploration does occur, a bias towards under-exploration in particular seems to be
-observed across a wide range of domains. In learned helplessness, for example, an organism
-experiences the absence of contol over the environment, learns that the
+In learned helplessness, for example, an organism
+experiences the absence of control over the environment, learns that the
 environment is uncontrollable, and thus ceases to take actions that might allow
@@
 as flying a plane or playing a sport \citep{Gopher1989}. In these settings, an ``emphasis
-change'' training method in which the focus of practice is repeatedly changed
-can lead to greater performance gains than unguided practice or more complex
-training methods. Emphasis change works because it encourages people to
-continually explore the performance space to find more effective cognitive and
-motor routines. Without these interventions, people often enter a ``local
+change'' training method that encourages people to continually explore the
+performance space leads to greater performance gains than unguided practice or more complex
+training methods.  Without this intervention, people often enter a ``local
 maximum'' in which exploration decreases and performance plateaus \citep{Yechiam2001}.
@@
 making, one might expect a clear pattern of under-exploration in lab
-experiments. Instead, however, lab results are mixed, with people sometimes
+experiments. Instead, lab results are mixed, with people sometimes
 under-exploring, sometimes over-exploring, and sometimes exploring close to an
@@
 other studies and paradigms
-\citep{tversky1966information, Busemeyer1988, hertwig2004decisions, Navarro2016, Juni2016, Sang2011}.
+\citep{tversky1966information, hertwig2004decisions, Navarro2016, Juni2016, Sang2011}.
 
@@
 of real-world exploratory choice, namely, the distribution of choices and
-outcomes over time, and that this spreading of exploratory choices over time may
-account for the tendency to under-explore.
+outcomes over time, and that because of people's bias towards immediate rewards,
+this spreading of choices over time may account for the tendency to under-explore.
 
@@
 underweighting of temporally distant rewards relative to close ones \citep{March1991, Levinthal1993}. 
-% \hl{does this someone undermine the novelty of your framing/prediction?}
-While
-choosing actions that are expected to immediately maximize reward is inherently
-exploitative, exploration often requires the decision-maker to suffer short-term
-costs in search of better future options.
+In expectation, exploitation yields the greatest reward for the present choice;
+exploration yields lower expected reward for the present choice, but raises the
+expected reward for all future choices due to the possibility of discovering
+a better option. Thus, if future choices and rewards are weighted less than
+immediate ones, the value of exploration is diminished.
 
-There are two distinct possible causes of myopia \citep{Bartels2015}. The first is that the decision-maker is
+% While
+% choosing actions that are expected to immediately maximize reward is inherently
+% exploitative, exploration often requires the decision-maker to suffer short-term
+% costs in search of better future options.
+
+\begin{figure*}
+\centering
+\includegraphics[width=\textwidth]{figures/machinetimeline.png}
+\caption{A schematic illustration of the experiment timeline. In all experiments
+  and conditions, six 30-second work periods passed between consecutive choices
+  with a given machine. \emph{a}: in Experiment 1 and the immediate condition of
+  Experiment 2, the outcome for a choice was revealed and occured immediately
+  after the choice was made. \emph{b}: in the delayed condition of Experiment 2 (right),
+  the outcome of a choice was revealed and occurred after a delay of four work
+  periods.}
+\label{fig:machinetimeline}
+\end{figure*}
+
+There are two distinct possible forms of myopia \citep{Bartels2015}. The first is that the decision-maker is
 unaware of or fails to consider the future. This form of myopia is likely one
@@
 patterns of exploration based on their expectations about future choices,
-showing that, at least to some degree they do consider the future \citep{Meyer1995, Wilson2014a, Rich2017}.
+showing that, at least to some degree they do consider the future \citep{Wilson2014a, Rich2017}.
 
@@
 A wealth of studies have shown that this form of myopia is an important factor
-in human decision making, and that people are strongly
+in human decision making, and that people are 
 biased towards  the present. For example, people will often prefer a larger, later monetary
@@
 Present bias has the clear potential to cause people to explore too little, by
-increasing the perceived immediate gains of exploiting relative to the delayed
-gains of exploring. However, there has been no experimental research
+increasing the weight of the greater expected immediate gains of exploiting relative to
+the greater expected delayed gains of exploring. However, there has been no experimental research
 examining how people's bias towards the present affects exploratory choice. In
@@
 rewards are generally not consumed on a trial-by-trial basis. For example, a participant
-might recieve ``points'' on each trial, which are converted to money when the
+might receive ``points'' on each trial, which are converted to money when the
 experiment ends. In this sort of paradigm, there is no reason to expect a bias
@@
 as predicted to well-studied variables. In the second, we introduce a
-delayed-outcome condition, which we predicted to increase exploration
-if people are present biased. However, we find no strong evidence that delaying outcomes affects
+delayed-outcome condition. If people are present biased, we predict this to
+increase exploration by decreasing the weight of the current choice's outcomes
+relative to future choice's outcomes.
+However, we find no strong evidence that delaying outcomes affects
 the rate of exploration. While this may support the conclusion that exploratory
@@
 consumption task would be unpleasant (moving sliders) or pleasant (watching a
-Youtube video). In Experiment 1, we sought to confirm that participants prefered
+YouTube video). In Experiment 1, we sought to confirm that participants preferred
 the video task to the slider task and would seek to increase video-watching
@@
 
-\begin{figure*}
-\centering
-\includegraphics[width=\textwidth]{figures/machinetimeline.png}
-\caption{A schematic illustration of the experiment timeline. In all experiments
-  and conditions, six 30-second work periods pass between consecutive choices
-  with a given machine. \emph{a}: in Experiment 1 and the immediate condition of
-  Experiment 2, the outcome for a choice is revealed and occurs immediately
-  after the choice is made. \emph{b}: in the delayed condition of Experiment 2 (right),
-  the outcome of a choice is revealed and occurs after a delay of four work
-  periods.}
-\label{fig:machinetimeline}
-\end{figure*}
-
 % \hl{Complex task to follow...  it might make sense to overview it and then
@@
 Participants were informed that their job was to perform a monotonous slider task that
-would be split into 30-second ``work periods'', but that they would be able to
+would be split into 30-second ``work periods,'' but that they would be able to
 make choices throughout the experiment that would give them a chance to watch a
-Youtube video instead. The number of remaining work periods in the experiment was shown
+YouTube video instead. The number of remaining work periods in the experiment was shown
 at the top of the screen, as was the number of seconds left in the current work period.
@@
 started at a random setting between 0 and 100, with the slider's value
-shown to its right, with a random horizontal offset so that the
+shown to its right, and with a random horizontal offset so that the
 sliders were not aligned. The participant's task was to use the mouse to move
@@
 Before beginning the experiment, participants chose one of four videos available
-on Youtube: an episode of ``Planet Earth'', and episode of ``The Great British
+on YouTube: an episode of ``Planet Earth'', and episode of ``The Great British
 Bakeoff'', and episode of ``Mythbusters'', or an Ellen Degeneres comedy special.
@@
 \includegraphics[width=\textwidth]{figures/machinemappictures.png}
-\caption{The machine display seen by participants to allow them to keep track of
-  the value of each of the six machines and the next time it would be ready to
+\caption{The machine display seen by participants. The display allowed participants track
+  the value of each machine and the next time each machine would be ready to
   make a choice or produce an outcome. Gray arrows have been added to depict the
@@
 \emph{Decision-making task.}
-Following the initial 10 periods, participants were shown a machine before each work period and had to select
-between two circular spinners with arrows at the top, as shown in Figure~\ref{fig:taskpictures}c: the ``current spinner''
+Following the initial 10 periods, participants were shown a machine before each
+work period and,
+as shown in Figure~\ref{fig:taskpictures}c, had to select
+between two circular spinners with arrows at the top: the ``current spinner''
 (exploit) and the ``new spinner'' (explore).
@@
 five black and five gold wedges. If a participant chose the
-current spinner, it spun (immediately or after four periods, depending
-on the condition) and, if the arrow landed on gold, the machine
+current spinner, it spun and, if the arrow landed on gold, the machine
 worked and the participant could watch the video. Initially, the
@@
 Finally, in order to induce exploration throughout the entire experiment, the
-six machines would occasionally ``reset'' after they ran. When this occured, a
+six machines would occasionally ``reset'' after they ran. When this occurred,
 the current spinner would be set to a new random value between 1/3 and 2/3 gold.
@@
 practice phases. First, they performed several trials of practice using the
-machines, with the actual work periods removed. Then, they performed three work
-periods practicing the slider task.
+machines, with the actual work periods removed. Then, they performed two work
+periods practicing the slider task and one work period practicing the video task.
 During the machine choices, participants had access to an ``info'' button at the
@@
 Participants were given a performance-based bonus of \$3.00 for completing the
-slider tasks accurately. If they missed fewer than 10\% of sliders throughout
+consumption tasks accurately. If they missed fewer than 10\% of sliders throughout
 the experiment and left the video paused less than 20\% of the time, they were
 not penalized. However, if they missed more sliders or left the video paused for
-longer, they lose 10 cents from their bonus for each additional percentage of
+longer, they lost 10 cents from their bonus for each additional percentage of
 sliders missed or time with the video paused. The running percentage of sliders
@@
 
-To analyze participants trial-by-trial decision-making, we conducted a hierarchical
+To analyze participants' trial-by-trial decision-making, we conducted a hierarchical
 Bayesian logistic regression using the Stan modeling language \citep{stan-software:2015}.
@@
 choices, we present individual-level model fits rather than the raw choice data.
-We find that at the population level, participants are less likely to choose a
-new spinner when the value of the current spinner is higher, with posterior
-probability $p>.999$. We also find that participants in the high-explore are more likely to explore a new spinner for a
+We found that at the population level, participants were less likely to choose a
+new spinner when the value of the current spinner was higher, with posterior
+probability $p>.999$. We also found that participants in the high-explore were more likely to explore a new spinner for a
 given current spinner value than those in the low-explore condition, with
@@
 similar to behavior in other exploratory choice tasks. Participants explored more
-when the current exploitative option is poor, as well as when the possible
-outcomes of exploration are good. While this experiment did not address whether the
+when the current exploitative option was poor, as well as when the possible
+payoffs of exploration were high. While this experiment did not address whether the
 consumption rewards produced a bias towards the present (a topic returned to in the
@@
 experiment, we introduce a delayed-outcome version of the task and predict that, if the
-consumption rewards induce a present bias, this version will increase exploration.
+consumption rewards induce a present bias, this version of the task will increase exploration.
 
@@
 ``process'' for four work periods, thus delaying
-the outcome by over 2 minutes Figure~\ref{fig:machinemappictures}b. The participant then returned to the
+the outcome by over 2 minutes (Figure~\ref{fig:machinemappictures}b). The participant then returned to the
 machine to observe its outcome and either perform the slider task or
@@
 was a temporal delay imposed between their choices and the received outcomes.
-However, there may be a small interaction between current spinner value and
+However, there may have been a small interaction between current spinner value and
 condition, such that participants in the delayed condition were less sensitive
@@
 
-In this paper we presented a novel task using immediately-consumed rewards that
-may be used to study present bias in the context of exploratory choice. In
+In this paper we presented a novel task using immediately-consumed rewards
+to study present bias in the context of exploratory choice. We predicted that
+adding a temporal delay between choices and outcomes would increase exploration
+by reducing the weight of the current choice's outcome (which is maximized
+through exploitation) relative to future choices' outcomes (which are increased
+through exploration). In
 Experiment 1 we showed that participants appeared to understand the task and
-respond reasonably to maximize their time in the video-watching task. They are
-more likely to explore when the current exploitable option (current spinner) has
-low value, and change their degree of exploration based on the range of possible
-outcomes from choosing a new spinner. However, in Experiment 2 we found that
+respond reasonably to maximize their time in the video-watching task. They were
+more likely to explore when the current exploitable option (current spinner) had
+low value, and changed their degree of exploration based on the range of possible
+outcomes from choosing a new spinner. However, in contrast to our prediction, in Experiment 2 we found that
 participants did not explore any more when there was a temporal delay between
-choices and outcomes, even though we expected this delay to reduce the present
-bias towards the greater immediate rewards of exploitation.
+choices and outcomes.
 
@@
 create a bias towards immediate video-watching in the time scale used in our
-experiments. To test this, we might use a simple ``concurrent schedule''
+experiments. To test this, we might use a simple 
 experiment in which two buttons produce video-watching time at for different
@@
 
-%%% this is a little too apologetic. who cares about our excitement! ;)
-%While introduces difficulties not found in standard decision-making tasks, we
-%remain excited about the possibility of creating a paradigm for the use of
-%immediately-consumable rewards in dynamic decision-making tasks. Such a paradigm
-%could allow not just a test of whether present bias affects exploratory choice,
-%but a greater focus on a variety of motivational factors often neglected by
-%cognitive science.
-
 \begin{small}
   \noindent
-\textbf{Acknowledgments} 
-\hl{FUNDING?}
+\textbf{Acknowledgments}  This research was supported by NSF grant BCS-1255538 and a John S. McDonnell Foundation Scholar Award to TMG.
 \end{small}
